---
title: "Supplementary results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=1cm
header-includes:
  - \usepackage{eso-pic,graphicx,transparent}
  - \usepackage{fancyhdr}
  - \usepackage{lastpage}
output:
  bookdown::pdf_document2: 
    toc_depth: 5
    toc_float: true
    number_sections: true
    keep_tex: true
  bookdown::html_document2: 
    toc: true
    toc_float: true
    keep_md: true
  md_document:
    variant: markdown_github
  bookdown::word_document2: default
  bookdown::github_document2:
    toc: true
    appendix: true
    fig_caption: yes
    number_sections: yes
    toc_float: true
    pandoc_args: --webtex
always_allow_html: yes 
params:
  lbfile: './data/20250612 updated scenario delivery and costing.xlsx'
---

\fancypagestyle{plain}{%
  \renewcommand{\headrulewidth}{0pt}%
  \fancyhf{}%
  \fancyfoot[R]{\footnotesize \thepage}
  \setlength\footskip{0pt}
}
\pagestyle{plain}
  

```{r setup, include=FALSE}
library(knitr) 
library(bookdown)
library(pander)    
library(data.table)
library(haven)
library(tidyr)
library(dplyr)
library(stringi)
library(ggplot2)  
library(RColorBrewer)
# library(splines)
library(viridis)
library(hrbrthemes)
# library(MASS)
library(evmix)
library(scales)
library(patchwork)
library(ggh4x)
library(purrr)


panderOptions('round',2)
panderOptions('table.split.table', Inf)

knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')

decimalplaces <- function(x) {
    if ((x %% 1) != 0) {
        nchar(strsplit(sub('0+$', '', as.character(format(x,scientific = F))), ".", fixed=TRUE)[[1]][[2]])
    } else {
        return(0)
    }
}

format_to_print2 <- function(x,z=2){
  v <- signif(x,z)
  sapply(v, function(i)  formatC(i, format="f", digits=decimalplaces(i), big.mark=","))
}

```

```{r scentab,echo=F,warning=F,message=F}

tab <- readxl::read_xlsx(params$lbfile)
strategies <- c('No Closures','School Closures','Economic Closures','Elimination')
newpolicies <- c('NC','RC1','RC2','RC3')
income_levels <- c('LLMIC','UMIC','HIC')
nScen = (ncol(tab)-2)/6
scenario_levels <- 1:nScen
scenario_names <- colnames(tab)[!grepl('^\\.',colnames(tab))]
scen_to_keep <- 1+c(0,1,4,5,8,9,12)
bau_scens = which(grepl('BAU',scenario_names))
scen_scens = which(!grepl('BAU',scenario_names))
bau_names = scenario_names[bau_scens]
# just_scen_names = scenario_names[scen_scens]

# nbscens = length(bau_names)
ncscens = length(scen_scens)

```















```{r resultsBAU,echo=F,warning=F,message=F}


result_cols <- c('Cost','YLL','School','GDP_loss')

allresults <- data.frame()

for (i in 1:length(income_levels)){
    income_level <- income_levels[i];
    inputtab <- read.csv(paste0('code/results/inputs_',income_level,'.csv'),header=T);
    for (k in 1:length(strategies)){
        inp3 <- strategies[k];
        results <- read.csv(paste0('code/results/outputs/outputs_',inp3,'_',income_level,'_scen1.csv'),header=T);
        results$gdp <- inputtab$gdp
        results$Policy <- inp3
        results$igroup <- income_level
        allresults <- rbind(allresults,results)
    }
}
nvar <- ncol(results) - 2
for(i in 1:nvar){
  allresults[,i] = as.numeric(allresults[,i]);
}
for(i in (nvar-3):nvar){
  allresults[,i] = allresults[,i]/allresults$gdp*100;
  # allresults[,i] = log(allresults[,i]);
}
vioplot <- reshape2::melt(allresults[,(nvar-3):ncol(allresults)],id.vars=c('Policy','igroup'))
vioplot$igroup <- factor(vioplot$igroup,levels=c('LLMIC','UMIC','HIC'))
vioplot$Policy <- factor(vioplot$Policy,levels=strategies, labels=newpolicies)
vioplot$variable <- factor(vioplot$variable,
                           levels=c('Cost','YLL','School','GDP_loss'),
                           labels=c('Total loss','YLL','Education','GDP'))


```

\newpage



```{r tabres0,echo=F,warning=F,message=F,results='asis'}

meltmeantab <- c()
allallresults <- tabresall <- data.frame()
nsamples <- nrow(results)
vsls <- gdps <- p <- topresults <- list()
for(sl in scenario_levels){
  scenario <- scenario_names[sl]
  p[[sl]] <- list()

  allresults <- data.frame()
  for (i in 1:length(income_levels)){
    income_level <- income_levels[i];
    inputtab <- read.csv(paste0('code/results/inputs_',income_level,'.csv'),header=T);
    vsls[[i]] <- inputtab$vsl * inputtab$gdp
    gdps[[i]] <- 1e6 * inputtab$gdp
    for (k in 1:length(strategies)){
      inp3 <- strategies[k];
      
      results <- read.csv(paste0('code/results/outputs/outputs_',inp3,'_',income_level,'_scen',sl,'.csv'),header=T);
      results$Policy <- newpolicies[k]
      results$igroup <- income_level
      results$samplei <- 1:nrow(results)
      # results <- cbind(results,inputtab)
      results$Costsl <- results$Cost/(inputtab$vsl * inputtab$gdp)
      results$Costpc <- 100 * results$Cost/inputtab$gdp
      results$gdp <- inputtab$gdp
      allresults <- rbind(allresults,results)
    }
  }
  nvar <- which(colnames(results)=='GDP_loss')
  outcomenames <- colnames(results)[ (nvar-3):nvar]
  for(i in 1:nvar){
    allresults[,i] = as.numeric(allresults[,i]);
  }
  for(i in (nvar-3):nvar){
    allresults[,i] = allresults[,i]/allresults$gdp*100;
    # allresults[,i] = log(allresults[,i]);
  }
  vioplot <- reshape2::melt(allresults[,c(outcomenames,'Policy','igroup')],id.vars=c('Policy','igroup'))
  vioplot$igroup <- factor(vioplot$igroup,levels=c('LLMIC','UMIC','HIC'))
  vioplot$Policy <- factor(vioplot$Policy,levels=newpolicies, labels=newpolicies)
  vioplot$variable <- factor(vioplot$variable,
                             levels=c('Cost','YLL','School','GDP_loss'),
                             labels=c('Total loss','YLL','Education','GDP'))
  
  p[[sl]] <- ggplot(vioplot,aes(fill=Policy,colour=Policy, y=value, x=Policy)) + 
    geom_violin(position="dodge", scale='width') +
    facet_grid(variable~igroup, scales="free") + 
      scale_fill_viridis(discrete=T, name="",option='inferno',end=.925,direction=-1) +
      scale_colour_viridis(discrete=T, name="",option='inferno',end=.925,direction=-1) +
      theme_bw(base_size = 12)  +
      xlab("") +
      ylab("Percent of GDP") +
    scale_x_discrete(labels=c('','','',''))+ 
    theme(axis.ticks.x=element_blank(),
          panel.spacing = unit(0.5, "lines"),
          legend.position = 'bottom',
          legend.margin=margin(t=-30))

  if(sl==1){
    var_names <- vioplot$variable %>% unique()
    facet_lims <- purrr::map(1:4, function(x){
      lims <- ggplot_build(p[[sl]])$layout$panel_scales_y[[x]]$range$range
      set_names(lims, c("Lower", "Upper"))
  }) %>%
      `names<-`(var_names) %>%
      bind_rows(.id = "Facet")
  }else{
    p[[sl]] <- p[[sl]] + ggh4x::facetted_pos_scales(y = list(
      variable == 'Total loss' ~ scale_y_continuous(limits = unlist(facet_lims[facet_lims$Facet=='Total loss',2:3])),
      variable == 'YLL' ~ scale_y_continuous(limits = unlist(facet_lims[facet_lims$Facet=='YLL',2:3])),
      variable == 'Education' ~ scale_y_continuous(limits = unlist(facet_lims[facet_lims$Facet=='Education',2:3])),
      variable == 'GDP' ~ scale_y_continuous(limits = unlist(facet_lims[facet_lims$Facet=='GDP',2:3]))
    ))
  }
  
  allresults$scenario <- scenario
  
  allallresults <- rbind(allallresults,allresults)
  
  setDT(allresults)
  allresults[,Costrnd:=Cost]
  allresults[Policy!='NC',Costrnd:=Cost*(1+runif(sum(allresults$Policy!='NC'))*1e-8)]
  allresults[,mincost:=Costrnd==min(Costrnd),by=.(igroup,samplei)]
  
  meantab <- allresults[,.(yll=mean(YLL),
                          education=mean(School),
                          gdp=mean(GDP_loss),
                          nmin=sum(mincost)),by=.(Policy,igroup)]
  meltmean <- reshape2::melt(meantab,id.vars=c('Policy','igroup','nmin'))
  meltmean$scen <- scenario
  meltmeantab <- rbind(meltmeantab,meltmean)
  
  topresults[[sl]] <- subset(allresults,mincost)
  setorder(topresults[[sl]],samplei)
  appendtopres <- copy(topresults[[sl]])
  appendtopres$Policy <- 'Loss minimising'
  allresults <- rbind(allresults,appendtopres)
  allresults[,iqr:=rank(Cost)%in%c(round(nsamples/4):round(3*nsamples/4)),by=.(Policy,igroup)]
  tabres <- allresults[iqr==T,list(costs=paste0(round(range(Cost),1),collapse='--\u200B'),
                          yll=paste0(round(mean(YLL/Cost*100),1),collapse='--\u200B'),
                          education=paste0(round(mean(School/Cost*100),1),collapse='--\u200B'),
                          gdp=paste0(round(mean(GDP_loss/Cost*100),1),collapse='--\u200B')
                          ),by=.(Policy,igroup)]
  colnames(tabres) <- c('Policy','Income group','Total loss','YLL','Education','GDP')
  tabres$Scenario <- scenario
  tabresall <- rbind(tabresall,tabres)
  

}

tabresall$Policy <- factor(tabresall$Policy,levels=c(newpolicies,'Loss minimising'))
tabresall$`Income group` <- factor(tabresall$`Income group`,levels=c(income_levels,'Loss minimising'))
setorder(tabresall,Scenario,`Income group`,Policy)
# tabresall[,Scenario:=paste0(`SARS-X vaccine`,' days, ',c('no BPSV','BPSV')[BPSV]),by=.(`SARS-X vaccine`,BPSV)]
tabresall <- cbind(tabresall[,7],tabresall[,-c(7)])
lg <- paste0('Results for all vaccination levels, scenarios and country types. Mean values and 95% prediction intervals. \\label{tab:tabres}')
# if (!knitr::is_html_output(excludes = "markdown")) { 
#   cat(pander(tabresall,caption=lg))
# }else{
#   x <- knitr::kable(tabresall,escape=F, "html",caption=lg)
#   (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
# }
```





# Results for all income levels and policy options

```{r tabresall,echo=F,warning=F,message=F,results='asis'}
lmtabresall <- subset(tabresall,Policy!='Loss minimising')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(lmtabresall,caption=lg))
}else{
  x <- knitr::kable(lmtabresall,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```



\newpage


# Results for all income levels and best policy option


```{r tabres,echo=F,warning=F,message=F,results='asis'}

lmtabresall <- subset(tabresall,Policy=='Loss minimising')
lmtabresall$Policy <- NULL
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(lmtabresall,caption=lg))
}else{
  x <- knitr::kable(lmtabresall,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```



\newpage

# Violin plots


```{r violintop,fig.cap='Model results for the cost-minimising responses.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}

alltopresults <- do.call(rbind,topresults)
vioplot <- reshape2::melt(alltopresults[,c(outcomenames,'igroup','scenario'),with=F],id.vars=c('igroup','scenario'))
vioplot$igroup <- factor(vioplot$igroup,levels=c('LLMIC','UMIC','HIC'))
# vioplot$Policy <- factor(vioplot$Policy,levels=newpolicies)
vioplot$variable <- factor(vioplot$variable,
                           levels=c('Cost','YLL','School','GDP_loss'),
                           labels=c('Total loss','YLL','Education','GDP'))

ggplot(subset(vioplot,scenario%in%scenario_names[scen_to_keep]),aes(y=value, x=scenario, fill=scenario,colour=scenario)) + 
  geom_violin(position="dodge", scale='width') +
  facet_grid(variable~igroup, scales="free") + 
    scale_fill_viridis(discrete=T, name="",option='inferno',end=.925,direction=-1) +
    scale_colour_viridis(discrete=T, name="",option='inferno',end=.925,direction=-1) +
    theme_bw(base_size = 12)  +
    xlab("") +
    ylab("Percent of GDP") +
  scale_x_discrete(labels=NULL)+ 
  theme(axis.ticks.x=element_blank(),
        panel.spacing = unit(0.5, "lines"),
        legend.position = 'bottom',
        legend.margin=margin(t=-30))
```


\newpage
## Scenario violin plots

```{r violinBAU,fig.cap='Model results for the 365-day SARS-X vaccine without the BPSV.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}
p[[scen_to_keep[1]]]

# ggsave(p[[1]][[1]] + theme_bw(base_size=15),filename='BAUviolinspill.png',width=9,height=6.5)
# 
# p[[1]][[1]]

```

```{r violinscen1,fig.cap='Model results for the 100-day SARS-X vaccine without the BPSV.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}
p[[scen_to_keep[2]]]
```

```{r violinscen2,fig.cap='Model results for the 365-day SARS-X vaccine with the BPSV.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}
p[[scen_to_keep[3]]]
```

```{r violinscen3,fig.cap='Model results for the 100-day SARS-X vaccine with the BPSV.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}
p[[scen_to_keep[4]]]
```

```{r violinscen4,fig.cap='Model results for the 100-day SARS-X vaccine with the BPSV.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}
p[[scen_to_keep[5]]]
```






\newpage

# Expected value by policy choice

```{r expectedvaluesbar,fig.cap='Expected values of model results.',echo=F,warning=F,message=F,fig.height=6,fig.width=6,eval=T}

# meltmeantab$scen <- factor((meltmeantab$scen), 
#                                     levels=c('365, No BPSV','365, BPSV',
#                                              '100, No BPSV','100, BPSV'))
setDT(meltmeantab)
# meltmeantab[,cost:=sum(value),by=.(scen,igroup,Policy)]
# meltmeantab[,mincost:=cost==min(cost),by=.(scen,igroup)]
meltmeantab[,mincost:=nmin==max(nmin),by=.(scen,igroup)]
ggplot(subset(meltmeantab,scen%in%scenario_names[scen_to_keep])) + 
      geom_bar(aes(x=factor(Policy,levels=newpolicies,labels=newpolicies),
                   y=value,fill=factor(variable,levels=c('education','yll','gdp'),labels=c('Education','YLL','GDP')),
                   alpha=I(as.numeric(mincost)*.5+.5)),stat='identity') + 
  facet_grid(factor(igroup,levels=c('LLMIC','UMIC','HIC'))~scen) + 
      theme_bw(base_size=15) +
        scale_fill_viridis(discrete=T, name="") +
      labs(x='',y='Loss, % of GDP',fill='Loss type') +
        theme(axis.text.x = element_text(angle = 45,  hjust=1,vjust=1),
              legend.position = 'top') 
    # ggsave(p,filename='figures/expectedcosts.pdf',height=10)
# subset(meltmeantab,igroup=='HIC'&scen=='100, BPSV')

# ggplot(subset(meltmeantab,grepl('Secondary',igroupcountry))) + 
#       geom_bar(aes(x=factor(Policy,levels=c('No Closures','School Closures','Economic Closures','Elimination')),
#                    y=value,fill=factor(variable,levels=c('education','yll','gdp'),labels=c('Education','YLL','GDP')),
#                    alpha=I(as.numeric(mincost)*.5+.5)),stat='identity') + 
#   facet_grid(factor(igroupcountry,labels=c('LLMIC','UMIC','HIC'))~factor(scen,levels=c('BAU','100 days','BPSV'))) + 
#       theme_bw(base_size=15) +
#         scale_fill_viridis(discrete=T, name="",option='inferno',end=.925) +
#       labs(x='',y='Loss, % of GDP',fill='Loss type') +
#         theme(axis.text.x = element_text(angle = 45,  hjust=1,vjust=1),
#               legend.position = 'right') -> p
# 
# ggsave(p,filename = 'expectedvaluesbar.png',height=6,width=8)


```





```{r resultsdiff,echo=F,warning=F,message=F}

difflist <- c()
alldiffs <- list()
for(sl in scen_scens){
  scenario <- scenario_names[sl]
  alldiffs[[length(alldiffs)+1]] <- list()
  for(refsl in bau_scens){
    refscen <- scenario_names[refsl]
    difftab <- as.data.frame(readRDS(paste0('code/results/differences/difftab_',refsl,'_',sl,'.Rds')))
    alldiffs[[length(alldiffs)]][[which(bau_scens==refsl)]] <- difftab
    difftab$Cost = difftab$Cost/difftab$gdp*100;
    difftab$scen <- paste0(refscen,' to ',scenario)
    difftab$fromscen <- refscen
    difftab$toscen <- scenario
    difflist <- rbind(difflist,difftab)
  }
}

```

\newpage

# Violin plot of difference


```{r diffviolin,fig.cap='Model results for losses saved by 100-day and BPS vaccines.',echo=F,warning=F,message=F,fig.height=4,fig.width=6}

vioplot <- reshape2::melt(difflist[,colnames(difflist)%in%c('Cost','igroup','toscen','fromscen')],id.vars=c('igroup','toscen','fromscen'))
vioplot$igroup <- factor(vioplot$igroup,levels=c('LLMIC','UMIC','HIC'))
# vioplot$variable <- factor(vioplot$variable,
#                            levels=c('Cost','YLL','School','GDP_loss'),
#                            labels=c('Total','YLL','Education','GDP'))

ggplot(subset(vioplot,toscen%in%scenario_names[scen_to_keep]&fromscen%in%bau_names),aes(fill=toscen, colour=toscen, y=value, x=toscen)) + 
    geom_violin(position="dodge",scale='width') +
  facet_grid(fromscen~igroup, scales="free") + 
    scale_fill_viridis(discrete=T, name="") +
    scale_colour_viridis(discrete=T, name="") +
    theme_bw(base_size = 15)  +
    xlab("") +
    ylab("Percent of GDP") +
  scale_x_discrete(labels=NULL)+ 
  theme(axis.ticks.x=element_blank(),
        panel.spacing = unit(0.5, "lines"),
        legend.position = 'bottom',
        legend.margin=margin(t=-20))-> p
  # ggsave(p,filename = 'figures/resultsdiff.pdf',width=12,height=10)
  p
  
  

```



\newpage

# Differences

```{r tabresdiff,tab.cap='Value of vaccines: expected difference between BAU and 100 days. Mean values and 95% prediction intervals.',echo=F,warning=F,message=F}

meantab <- setDT(difflist)[,.(yll=mean(YLL),
                        education=mean(School),
                        gdp=mean(GDP_loss)),by=.(igroup,fromscen,toscen)]

difflist[,IQR:=rank(Cost)%in%c(round(nsamples/4):round(3*nsamples/4)),by=.(fromscen,toscen,igroup)]
tabres <- setDT(difflist)[IQR==T,list(costs=paste0(signif(range(Cost),2),collapse='--\u200B')
                        ),by=.(fromscen,toscen,igroup)]
# tabres <- dcast(tabres,formula=igroup+fromscen~toscen,value.var = 'costs',fill = '')

tabres[,igroup:=factor(igroup,levels=c('LLMIC','UMIC','HIC'))]
setorder(tabres,igroup,fromscen)
colnames(tabres)[1:3] <- c('From scenario','To scenario','Income group')


# write.csv(tabres,'cepi_results/Delta_LIR_IQR_IL.csv',row.names = F, quote = F)

# subtabres <- subset(tabres,`From scenario`%in%bau_names)[,!colnames(tabres)%in%scenario_names[-c(scen_to_keep)],with=F]

lg <- 'Value of vaccines as a percent of GDP: differences between BAU and scenarios. 75% uncertainty intervals. \\label{tab:tabresdiff}'
if (!knitr::is_html_output(excludes = "markdown")) { 
  pander(tabres,caption=lg)
}else{
    x <- knitr::kable(tabres,escape=F, "html",caption=lg)
  kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position")
}
```

\newpage

# Best-case scenarios


```{r bestcase,tab.cap='Value of vaccines',echo=F,warning=F,message=F}

plotlist <- readRDS('code/results/exXbars.Rds')
valuetablist <- readRDS('code/results/exXvalues.Rds')
valuetab <- do.call(rbind,valuetablist[[1]])
valuetab2 <- valuetablist[[2]]

colnames(valuetab)[2] = 'Scenario'
countercol = c(rep(rownames(valuetab2)[1:16],each=ncscens))

valuetabcounter <- setDT(subset(valuetab,counter=='counter'))
valuetabcounter$Counterfactual = countercol
rownames(valuetabcounter) <- NULL # rownames(valuetab2)[1:16]
valuetabcounter$`From scenario` <- NULL
valuetabcounter[,lq:=as.numeric(strsplit(Value,'--')[[1]][1]),by=Value]
valuetabcounter[,uq:=as.numeric(strsplit(Value,'--')[[1]][2]),by=Value]
valuetabcounter[,ml:=max(lq),by=Counterfactual]
valuetabcounter[,mu:=max(uq),by=Counterfactual]

vtab = subset(valuetabcounter,ml==lq|mu==uq)[,.(Scenario,Value,Counterfactual)]
bsc <- data.frame(sapply(unique(vtab$Counterfactual),function(x)with(subset(vtab,Counterfactual==x),paste0(Scenario,collapse=', '))))
colnames(bsc) <- 'Best-case scenarios'
bsc$Counterfactual <- rownames(bsc)
rownames(bsc) <- NULL
bsc <- bsc[,2:1]
lg <- paste0('Best-case scenarios: scenarios that have the highest $\\Delta$LIR lower quartile or upper quartile for each counterfactual. \\label{tab:bsc}')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(bsc,caption=lg))
}else{
  x <- knitr::kable(bsc,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```




\newpage

# Absolute values: LIR and numbers of deaths




```{r absvals,tab.cap='Absolute losses and numbers of deaths of one-in-thirty-years and one-in-one-hundred-years events under the BAU scenario.',echo=F,warning=F,message=F}

valuetabbau1 <- valuetab2[grepl('BAU1',rownames(valuetab2)),]
rownames(valuetabbau1) <- gsub(', BAU1','',rownames(valuetabbau1))

lg <- paste0('Absolute losses and numbers of deaths of one-in-X-years events under the BAU scenario. \\label{tab:absvals}')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(valuetabbau1,caption=lg))
}else{
  x <- knitr::kable(valuetabbau1,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```



\newpage

# Percent of GDP


```{r ex100gdp,tab.cap='Value of vaccines',echo=F,warning=F,message=F}

valuetabgdp <- setDT(subset(valuetab,counter=='gdp'))
valuetabgdp[,counter:=NULL]
valuetabgdp$Counterfactual = countercol
rownames(valuetabgdp) <- NULL # rownames(valuetab2)[1:16]
valuetabgdp$`From scenario` <- NULL
# subvaluetab <- valuetabgdp[,!colnames(valuetabgdp)%in%scenario_names[-c(scen_to_keep)]]
lg <- paste0('Value of vaccine investments in the event of a one-in-X-years pandemic (percent of global GDP). \\label{tab:ex100gdp}')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(valuetabgdp,caption=lg))
}else{
  x <- knitr::kable(valuetabgdp,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```



\newpage

# Percent of counterfactual


```{r ex100,tab.cap='Value of vaccines',echo=F,warning=F,message=F}
valuetabcounter[,counter:=NULL]
valuetabcounter[,lq:=NULL]
valuetabcounter[,uq:=NULL]
valuetabcounter[,ml:=NULL]
valuetabcounter[,mu:=NULL]

# subvaluetab <- valuetabcounter[,!colnames(valuetabcounter)%in%scenario_names[-c(scen_to_keep)]]
lg <- paste0('Value of vaccine investments in the event of a one-in-X-years pandemic (percent of counterfactual). \\label{tab:ex100}')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(valuetabcounter,caption=lg))
}else{
  x <- knitr::kable(valuetabcounter,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}
```





\newpage

# Distributions of loss

```{r ex30bar,fig.cap='Distributions of values of investments for once-in-thirty-years events.',echo=F,warning=F,message=F,fig.height=8,fig.width=10}
print(plotlist)
```

\newpage

# Expected values


```{r expvalue,tab.cap='Expected value of investments',echo=F,warning=F,message=F,results='asis'}

expvals1 <- readRDS('code/results/expvals_BAU1.Rds')
expvals2 <- readRDS('code/results/expvals_BAU2.Rds')
expvals <- expvals1[expvals2,,on='Scenario']
  #rbind(readRDS('code/results/expvals_BAU1.Rds'),readRDS('code/results/expvals_BAU2.Rds'))
subexp <- expvals#[,!colnames(expvals)%in%scenario_names[-c(scen_to_keep)],with=F]

lg <- paste0('Value of investments. \\label{tab:expval}')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(subexp,caption=lg))
}else{
  # (knitr::kable(subexp,escape=F, "html",caption=lg))
  x <- knitr::kable(subexp,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```


\newpage

# Policy choices

```{r choicestab,fig.cap='Frequency of mitigation policy choices.',echo=F,warning=F,message=F,fig.height=4,fig.width=6}

choicestab <- readRDS(paste0('code/results/choicestab.Rds'))
nsamples <- sum(choicestab$N)/3/nScen
ggplot(subset(choicestab,scenario%in%scenario_names[scen_to_keep])) +  
  geom_bar(aes(x=scenario,
               y=N/nsamples*100,
               fill=factor(policy,levels=strategies,labels=newpolicies)),stat='identity') + 
  facet_wrap(~factor(igroup,levels=c('LLMIC','UMIC','HIC'))) + 
  theme_bw(base_size=15) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_viridis(discrete=T, name="",option='inferno',end=.925,direction=-1) +
  labs(x='',y='Chosen policy, % of samples',fill='') +
  theme(axis.text.x = element_text(angle = 45,  hjust=1,vjust=1),
        legend.position = 'left') 

```


```{r kept,echo=F,warning=F,message=F,eval=T}

setDT(allallresults)
allallresults[,Costrnd:=Cost]
allallresults[Policy!='NC',Costrnd:=Cost*(1+runif(sum(allallresults$Policy!='NC'))*1e-8)]
allallresults[,mincost:=min(Costrnd),by=.(igroup,scenario,samplei)]
# allallresults[,mincost:=min(Cost),by=.(igroup,scenario,samplei)]
allallresults[,keeprow:=Costrnd==mincost]
allallresults$Policy <- factor(allallresults$Policy,levels=newpolicies,labels=newpolicies)

keptresults <- subset(allallresults,keeprow==T)

```


\newpage

## Policy choices vs. severity


```{r intensity,fig.cap='The relationship between policy choice and severity (total deaths).',echo=F,warning=F,message=F,fig.height=5,fig.width=6,eval=T}

keptresults[,qdeath:=rank(Deaths)/.N,by=.(igroup,scenario)]

ggplot(subset(keptresults,scenario%in%scenario_names[scen_to_keep])) + 
  geom_histogram(aes(x=qdeath,y=stat(count / sum(count)),fill=Policy,group=Policy),position='fill',bins = 10) +
  facet_grid(factor(igroup,levels=c('LLMIC','UMIC','HIC'))~scenario) + 
  theme_bw(base_size = 11) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_viridis_d(name="",option='inferno',end=.925,direction=-1) +
  theme(axis.text.x = element_text(angle = 45,  hjust=1,vjust=1),
        legend.position = 'top') + labs(fill='',x='Deaths per thousand (quantile)',y='Proportion')

```





