---
title: "Supplementary results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=1cm
header-includes:
  - \usepackage{eso-pic,graphicx,transparent}
  - \usepackage{fancyhdr}
  - \usepackage{lastpage}
output:
  bookdown::pdf_document2: 
    toc_depth: 5
    toc_float: true
    number_sections: true
    keep_tex: true
  bookdown::html_document2: 
    toc: true
    toc_float: true
    keep_md: true
  md_document:
    variant: markdown_github
  bookdown::word_document2: default
  bookdown::github_document2:
    toc: true
    appendix: true
    fig_caption: yes
    number_sections: yes
    toc_float: true
    pandoc_args: --webtex
always_allow_html: yes 
params:
  lbfile: './data/20251209 Vx manufacturing and delivery.xlsm'
---

\fancypagestyle{plain}{%
  \renewcommand{\headrulewidth}{0pt}%
  \fancyhf{}%
  \fancyfoot[R]{\footnotesize \thepage}
  \setlength\footskip{0pt}
}
\pagestyle{plain}
  

```{r setup, include=FALSE}
library(knitr) 
library(bookdown)
library(pander)    
library(data.table)
library(haven)
library(tidyr)
library(dplyr)
library(stringi)
library(ggplot2)  
library(RColorBrewer)
# library(splines)
library(viridis)
library(hrbrthemes)
# library(MASS)
library(evmix)
library(scales)
library(patchwork)
library(ggh4x)
library(purrr)


panderOptions('round',2)
panderOptions('table.split.table', Inf)

knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, results='asis')

decimalplaces <- function(x) {
    if ((x %% 1) != 0) {
        nchar(strsplit(sub('0+$', '', as.character(format(x,scientific = F))), ".", fixed=TRUE)[[1]][[2]])
    } else {
        return(0)
    }
}

format_to_print2 <- function(x,z=2){
  v <- signif(x,z)
  sapply(v, function(i)  formatC(i, format="f", digits=decimalplaces(i), big.mark=","))
}

```

```{r scentab,echo=F,warning=F,message=F}

tab <- readxl::read_xlsx(params$lbfile, sheet = 'Vx timeline')
colnames(tab)[colnames(tab)=='BAU1'] = 'BAU'
strategies <- c('No Closures','School Closures','Economic Closures','Elimination')
newpolicies <- c('NC','RC1','RC2','RC3')
income_levels <- c('LLMIC','UMIC','HIC')
nScen = (ncol(tab)-2)/6
scenario_levels <- 1:nScen
scenario_names <- colnames(tab)[!grepl('^\\.',colnames(tab))]
scen_to_keep <- 1+c(0,1,4,5,8,9,12)
bau_scens = which(grepl('BAU',scenario_names))
scen_scens = which(!grepl('BAU',scenario_names))
bau_names = scenario_names[bau_scens]
# just_scen_names = scenario_names[scen_scens]

# nbscens = length(bau_names)
ncscens = length(scen_scens)

```


```{r costall,echo=F,warning=F,message=F}

get_to_print = function(qu){
  if(abs(qu[1]-qu[2])<0.01*abs(qu[1])){
    format_to_print2(qu[1])
  }else{
    paste0(format_to_print2(qu), collapse='--\\hspace{0pt}')
  }
}
quants = c(1,3)/4
scenario_results = readRDS(file = 'code/results/all_cost_samples.Rds')
iqrs = list()
for(s in 1:nScen){
  costtab =   as.data.frame(do.call(cbind,lapply(scenario_results[[s]]$costs,function(x) do.call(cbind,x))))
  costtab$upfront_bpsv <- costtab$bpsv_rd_undiscounted <- NULL
  iqrs[[s]] = apply(costtab,2,function(x) get_to_print(quantile(x,quants)))

}
printtab = do.call(cbind,iqrs)
colnames(printtab) = scenario_names
newrownames = c(enabling='Enabling activities',
                dis_upfront_bpsv='BPSV upfront procurement',
                bpsv_rd_discounted='BPSV upfront R\\&D',
                totalupfront='Total upfront',
                capres='Reserved capacity cost',
                investigational_reserve='BPSV investigational reserve',
                totalannual='Total annual',
                ssv_rd='SSV R\\&D',
                ssv_proc_undiscounted='SSV procurement',
                ssv_delivery_undiscounted='SSV delivery',
                ssv_proc_discounted='SSV procurement',
                ssv_delivery_discounted='SSV delivery',
                bpsv_response_rd='BPSV response R\\&D',
                bpsv_proc='BPSV response procurement',
                bpsv_delivery='BPSV delivery',
                totaldiscounted='Total response',
                totalundiscounted='Total response')
printtab = cbind(newrownames[rownames(printtab)],printtab)
rownames(printtab) = NULL
  
lg = 'Costs.'
if (!knitr::is_html_output(excludes = "markdown")) {
  cat(pander(printtab,caption=lg))
}else{
  x <- knitr::kable(printtab,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```

```{r expval}
expvals1 <- readRDS('code/results/expvals_BAU.Rds')
resulttab = data.frame(Scenario=expvals1$Scenario, Upfront=NA, Annual=NA, Response=NA, LIR=expvals1$BAU)

for(s in 1:nScen){ #c(1,10)){# 
  
  thisscen = scenario_results[[s]]
  allupfront = with(thisscen$costs$upfront, enabling + dis_upfront_bpsv + bpsv_rd_discounted)
  allannual = with(thisscen$costs$annual, capres + investigational_reserve)
  allresp = with(thisscen$costs$response, ssv_rd + ssv_proc_undiscounted + ssv_delivery_undiscounted + 
                   bpsv_response_rd + bpsv_proc + bpsv_delivery)
  
  if(s==1){
    bauallupfront = allupfront
    bauallannual = allannual
    bauallresp = allresp
    
  }else{
    diffallupfront = allupfront - bauallupfront
    diffallannual = allannual - bauallannual
    diffallresp = allresp - bauallresp
    
    resulttab$Upfront[s-1] = get_to_print(quantile(diffallupfront,quants))
    resulttab$Annual[s-1] = get_to_print(quantile(diffallannual,quants))
    resulttab$Response[s-1] = get_to_print(quantile(diffallresp,quants))
    
  }
  
}

lg = 'All scenario results compared to BAU.'
if (!knitr::is_html_output(excludes = "markdown")) {
  cat(pander(resulttab,caption=lg))
}else{
  x <- knitr::kable(resulttab,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```

```{r scenarios,message=F,echo=F,warning=F,fig.cap='Vaccine coverage over time by income level.',fig.height=4}

tab[1,c(is.na(tab[1,]))] = ''

df_list <- list()
for(scen in 1:nScen){
  for(il in 1:3){
    start_index <- (scen-1)*6 + 1 + 2
    # print(any(tab[-c(1,2),start_index]>0 & tab[-c(1,2),start_index + 3]>0))
    get_index <- start_index - 1 + il
    ilevel <- gsub('s','',tab[2,get_index])
    scenario <- colnames(tab)[start_index]
    vals <- 0
    for(i in c(0,3)){
      vaccine <- tab[1,start_index+i][[1]]
      vindex <- get_index + i
      vals <- vals +  as.numeric(tab[-c(1:2),vindex][[1]])
    }
    df_list[[length(df_list)+1]] <- data.frame(Day=1:length(vals), Percent=cumsum(as.numeric(vals))*100, #Vaccine=vaccine,
                                               Scenario=scenario, Income=ilevel)
  }
}

plotscens <- do.call(rbind,df_list)
plotscens$Week = plotscens$Day/7
# plotscens$Percent[plotscens$Percent>80] <- 80
# plotscens$Percent[plotscens$Percent>25&plotscens$Vaccine=='BPSV'] <- 25
pl1 = ggplot(subset(plotscens,!Scenario%in%c('BAU2'))) + 
  geom_line(aes(x=Week,y=Percent,colour=factor(Scenario,levels=scenario_names[c(2:11,1)])),linewidth=1,alpha=1) +
  facet_grid(~factor(Income,levels=c('LLMIC','UMIC','HIC')),scales='free_y') + 
  theme_bw(base_size=15) + labs(colour='') +
  theme(legend.position = 'top') + 
  guides(colour=guide_legend(nrow=2,byrow=F))

pl2 = pl1 +
  scale_colour_manual(values = with(subset(plotscens, Scenario != "BAU2"), {
    lev <- levels(factor(Scenario))
    setNames(c("black", colorspace::qualitative_hcl(length(lev) - 1, c = 100, l = 65)),
             c("BAU", setdiff(lev, "BAU")))
  }))

pl1
pl2
# ggplot(subset(plotscens,Income=='HIC'&Scenario%in%c('S12'))) + 
#   geom_line(aes(x=Day,y=c(0,diff(Percent)),colour=Scenario),linewidth=1,alpha=.75) +
#   theme_bw(base_size=15) + labs(colour='') +
#   theme(legend.position = 'top') + 
#   guides(colour=guide_legend(nrow=2,byrow=F))
```


```{r costresultsbau1,echo=F,warning=F,message=F, results='asis'}
lbfileold = './data/20251104 updated scenario delivery and costing.xlsx'
costresults <- readxl::read_xlsx(lbfileold,sheet = 3)
costresults$`Cost Category`[which(costresults$`Cost Category`=='Difference to BAU')+0:4] <- 'Difference to BAU'
costresults <- costresults[!is.na(costresults$`Cost Category`),]
costresults <- costresults[costresults$`Cost Category`!='Cost Category',]
newClabels <- c('Mean','C5','C25','C50','C75','C95')
oldClabels <- unique(costresults[[2]])
costresults[[2]] <- newClabels[match(costresults[[2]],oldClabels)]
for(i in 3:ncol(costresults)) {
  costresults[[i]] <- gsub('\\.','{\\cdotp}',format_to_print2(as.numeric(costresults[[i]]),2))
}
boldrows = grepl('Total',costresults$`Cost Category`) &! grepl('Cashflow',costresults$`Cost Category`)
for(i in 1:2) {
  costresults[[i]][boldrows] <- paste0('\\textbf{',costresults[[i]][boldrows],'}')
}
for(i in 3:ncol(costresults)) {
  costresults[[i]][boldrows] <- paste0('\\mathbf{',costresults[[i]][boldrows],'}')
}
for(i in 3:ncol(costresults)) {
  costresults[[i]] <- paste0('$',costresults[[i]],'$')
}
lg = 'Full results from costing model for BAU. Values given are the arithmetic mean, and centiles 5, 25, 50, 75 and 95. Units are million USD. Results are reported to two signficant figures.'
# if (!knitr::is_html_output(excludes = "markdown")) {
#   cat(pander(costresults,caption=lg))
# }else{
#   x <- knitr::kable(costresults,escape=F, "html",caption=lg)
#   (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
# }

```


```{r costresultsbau2,echo=F,warning=F,message=F,eval=F}
costresults <- readxl::read_xlsx(lbfileold,sheet = 4)
costresults$`Cost Category`[which(costresults$`Cost Category`=='Difference to BAU')+0:4] <- 'Difference to BAU'
costresults <- costresults[!is.na(costresults$`Cost Category`),]
costresults <- costresults[costresults$`Cost Category`!='Cost Category',]
newClabels <- c('Mean','C5','C25','C50','C75','C95')
oldClabels <- unique(costresults[[2]])
costresults[[2]] <- newClabels[match(costresults[[2]],oldClabels)]
for(i in 3:ncol(costresults)) {
  costresults[[i]] <- gsub('\\.','{\\cdotp}',format_to_print2(as.numeric(costresults[[i]]),2))
}
boldrows = grepl('Total',costresults$`Cost Category`) &! grepl('Cashflow',costresults$`Cost Category`)
for(i in 1:2) {
  costresults[[i]][boldrows] <- paste0('\\textbf{',costresults[[i]][boldrows],'}')
}
for(i in 3:ncol(costresults)) {
  costresults[[i]][boldrows] <- paste0('\\mathbf{',costresults[[i]][boldrows],'}')
}
for(i in 3:ncol(costresults)) {
  costresults[[i]] <- paste0('$',costresults[[i]],'$')
}
lg = 'Full results from costing model for BAU2. Values given are the arithmetic mean, and centiles 5, 25, 50, 75 and 95. Units are million USD. Results are reported to two signficant figures.'
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(costresults,caption=lg))
}else{
  x <- knitr::kable(costresults,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```














```{r resultsBAU,echo=F,warning=F,message=F}


result_cols <- c('Cost','YLL','School','GDP_loss')

allresults <- data.frame()

for (i in 1:length(income_levels)){
    income_level <- income_levels[i];
    inputtab <- read.csv(paste0('code/results/inputs_',income_level,'.csv'),header=T);
    for (k in 1:length(strategies)){
        inp3 <- strategies[k];
        results <- read.csv(paste0('code/results/outputs/outputs_',inp3,'_',income_level,'_scen1.csv'),header=T);
        results$gdp <- inputtab$gdp
        results$Policy <- inp3
        results$igroup <- income_level
        allresults <- rbind(allresults,results)
    }
}
nvar <- ncol(results) - 2
for(i in 1:nvar){
  allresults[,i] = as.numeric(allresults[,i]);
}
for(i in (nvar-3):nvar){
  allresults[,i] = allresults[,i]/allresults$gdp*100;
  # allresults[,i] = log(allresults[,i]);
}
vioplot <- reshape2::melt(allresults[,(nvar-3):ncol(allresults)],id.vars=c('Policy','igroup'))
vioplot$igroup <- factor(vioplot$igroup,levels=c('LLMIC','UMIC','HIC'))
vioplot$Policy <- factor(vioplot$Policy,levels=strategies, labels=newpolicies)
vioplot$variable <- factor(vioplot$variable,
                           levels=c('Cost','YLL','School','GDP_loss'),
                           labels=c('Total loss','YLL','Education','GDP'))


```

\newpage



```{r tabres0,echo=F,warning=F,message=F,results='asis'}

meltmeantab <- c()
allallresults <- tabresall <- data.frame()
nsamples <- nrow(results)
vsls <- gdps <- p <- topresults <- list()
for(sl in scenario_levels){
  scenario <- scenario_names[sl]
  p[[sl]] <- list()

  allresults <- data.frame()
  for (i in 1:length(income_levels)){
    income_level <- income_levels[i];
    inputtab <- read.csv(paste0('code/results/inputs_',income_level,'.csv'),header=T);
    vsls[[i]] <- inputtab$vsl * inputtab$gdp
    gdps[[i]] <- 1e6 * inputtab$gdp
    for (k in 1:length(strategies)){
      inp3 <- strategies[k];
      
      results <- read.csv(paste0('code/results/outputs/outputs_',inp3,'_',income_level,'_scen',sl,'.csv'),header=T);
      results$Policy <- newpolicies[k]
      results$igroup <- income_level
      results$samplei <- 1:nrow(results)
      # results <- cbind(results,inputtab)
      results$Costsl <- results$Cost/(inputtab$vsl * inputtab$gdp)
      results$Costpc <- 100 * results$Cost/inputtab$gdp
      results$gdp <- inputtab$gdp
      allresults <- rbind(allresults,results)
    }
  }
  nvar <- which(colnames(results)=='GDP_loss')
  outcomenames <- colnames(results)[ (nvar-3):nvar]
  for(i in 1:nvar){
    allresults[,i] = as.numeric(allresults[,i]);
  }
  for(i in (nvar-3):nvar){
    allresults[,i] = allresults[,i]/allresults$gdp*100;
    # allresults[,i] = log(allresults[,i]);
  }
  vioplot <- reshape2::melt(allresults[,c(outcomenames,'Policy','igroup')],id.vars=c('Policy','igroup'))
  vioplot$igroup <- factor(vioplot$igroup,levels=c('LLMIC','UMIC','HIC'))
  vioplot$Policy <- factor(vioplot$Policy,levels=newpolicies, labels=newpolicies)
  vioplot$variable <- factor(vioplot$variable,
                             levels=c('Cost','YLL','School','GDP_loss'),
                             labels=c('Total loss','YLL','Education','GDP'))
  
  p[[sl]] <- ggplot(vioplot,aes(fill=Policy,colour=Policy, y=value, x=Policy)) + 
    geom_violin(position="dodge", scale='width') +
    facet_grid(variable~igroup, scales="free") + 
      scale_fill_viridis(discrete=T, name="",option='inferno',end=.925,direction=-1) +
      scale_colour_viridis(discrete=T, name="",option='inferno',end=.925,direction=-1) +
      theme_bw(base_size = 12)  +
      xlab("") +
      ylab("Percent of GDP") +
    scale_x_discrete(labels=c('','','',''))+ 
    theme(axis.ticks.x=element_blank(),
          panel.spacing = unit(0.5, "lines"),
          legend.position = 'bottom',
          legend.margin=margin(t=-30))

  if(sl==1){
    var_names <- vioplot$variable %>% unique()
    facet_lims <- purrr::map(1:4, function(x){
      lims <- ggplot_build(p[[sl]])$layout$panel_scales_y[[x]]$range$range
      set_names(lims, c("Lower", "Upper"))
  }) %>%
      `names<-`(var_names) %>%
      bind_rows(.id = "Facet")
  }else{
    p[[sl]] <- p[[sl]] + ggh4x::facetted_pos_scales(y = list(
      variable == 'Total loss' ~ scale_y_continuous(limits = unlist(facet_lims[facet_lims$Facet=='Total loss',2:3])),
      variable == 'YLL' ~ scale_y_continuous(limits = unlist(facet_lims[facet_lims$Facet=='YLL',2:3])),
      variable == 'Education' ~ scale_y_continuous(limits = unlist(facet_lims[facet_lims$Facet=='Education',2:3])),
      variable == 'GDP' ~ scale_y_continuous(limits = unlist(facet_lims[facet_lims$Facet=='GDP',2:3]))
    ))
  }
  
  allresults$scenario <- scenario
  
  allallresults <- rbind(allallresults,allresults)
  
  setDT(allresults)
  allresults[,Costrnd:=Cost]
  allresults[Policy!='NC',Costrnd:=Cost*(1+runif(sum(allresults$Policy!='NC'))*1e-8)]
  allresults[,mincost:=Costrnd==min(Costrnd),by=.(igroup,samplei)]
  
  meantab <- allresults[,.(yll=mean(YLL),
                          education=mean(School),
                          gdp=mean(GDP_loss),
                          nmin=sum(mincost)),by=.(Policy,igroup)]
  meltmean <- reshape2::melt(meantab,id.vars=c('Policy','igroup','nmin'))
  meltmean$scen <- scenario
  meltmeantab <- rbind(meltmeantab,meltmean)
  
  topresults[[sl]] <- subset(allresults,mincost)
  setorder(topresults[[sl]],samplei)
  appendtopres <- copy(topresults[[sl]])
  appendtopres$Policy <- 'Loss minimising'
  allresults <- rbind(allresults,appendtopres)
  allresults[,iqr:=rank(Cost)%in%c(round(nsamples/4):round(3*nsamples/4)),by=.(Policy,igroup)]
  tabres <- allresults[iqr==T,list(costs=paste0(round(range(Cost),1),collapse='--\u200B'),
                          yll=paste0(round(mean(YLL/Cost*100),1),collapse='--\u200B'),
                          education=paste0(round(mean(School/Cost*100),1),collapse='--\u200B'),
                          gdp=paste0(round(mean(GDP_loss/Cost*100),1),collapse='--\u200B')
                          ),by=.(Policy,igroup)]
  colnames(tabres) <- c('Policy','Income group','Total loss','YLL','Education','GDP')
  tabres$Scenario <- scenario
  tabresall <- rbind(tabresall,tabres)
  

}

tabresall$Policy <- factor(tabresall$Policy,levels=c(newpolicies,'Loss minimising'))
tabresall$`Income group` <- factor(tabresall$`Income group`,levels=c(income_levels,'Loss minimising'))
setorder(tabresall,Scenario,`Income group`,Policy)
# tabresall[,Scenario:=paste0(`SARS-X vaccine`,' days, ',c('no BPSV','BPSV')[BPSV]),by=.(`SARS-X vaccine`,BPSV)]
tabresall <- cbind(tabresall[,7],tabresall[,-c(7)])
lg <- paste0('Results for all policies, scenarios and country types. Mean values and 95% prediction intervals. \\label{tab:tabresall}')
# if (!knitr::is_html_output(excludes = "markdown")) { 
#   cat(pander(tabresall,caption=lg))
# }else{
#   x <- knitr::kable(tabresall,escape=F, "html",caption=lg)
#   (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
# }
```





# Results for all income levels and policy options

```{r tabresall,echo=F,warning=F,message=F,results='asis'}

# for(i in 4:7) tabresall[[i]] <- gsub('\\.','·',tabresall[[i]])

lmtabresall <- subset(tabresall,Policy!='Loss minimising'&Scenario!='BAU2')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(lmtabresall,caption=lg))
}else{
  x <- knitr::kable(lmtabresall,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```



\newpage


# Results for all income levels and best policy option


```{r tabres,echo=F,warning=F,message=F,results='asis'}

lg <- paste0('Best-policy results for all scenarios and country types. Mean values and 95% prediction intervals. \\label{tab:tabres}')

lmtabresall <- subset(tabresall,Policy=='Loss minimising'&Scenario!='BAU2')
lmtabresall$Policy <- NULL
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(lmtabresall,caption=lg))
}else{
  x <- knitr::kable(lmtabresall,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```



\newpage

# Violin plots


```{r violintop,fig.cap='Model results for the cost-minimising responses.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}

alltopresults <- do.call(rbind,topresults)
vioplot <- reshape2::melt(alltopresults[,c(outcomenames,'igroup','scenario'),with=F],id.vars=c('igroup','scenario'))
vioplot$igroup <- factor(vioplot$igroup,levels=c('LLMIC','UMIC','HIC'))
# vioplot$Policy <- factor(vioplot$Policy,levels=newpolicies)
vioplot$variable <- factor(vioplot$variable,
                           levels=c('Cost','YLL','School','GDP_loss'),
                           labels=c('Total loss','YLL','Education','GDP'))

ggplot(subset(vioplot,scenario%in%scenario_names[scen_to_keep]),aes(y=value, x=scenario, fill=scenario,colour=scenario)) + 
  geom_violin(position="dodge", scale='width') +
  facet_grid(variable~igroup, scales="free") + 
    scale_fill_viridis(discrete=T, name="",option='inferno',end=.925,direction=-1) +
    scale_colour_viridis(discrete=T, name="",option='inferno',end=.925,direction=-1) +
    theme_bw(base_size = 12)  +
    xlab("") +
    ylab("Percent of GDP") +
  scale_x_discrete(labels=NULL)+ 
  theme(axis.ticks.x=element_blank(),
        panel.spacing = unit(0.5, "lines"),
        legend.position = 'bottom',
        legend.margin=margin(t=-30))
```


\newpage
## Scenario violin plots

```{r violinBAU,fig.cap='Model results for the 365-day SARS-X vaccine without the BPSV.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}
p[[scen_to_keep[1]]]

# ggsave(p[[1]][[1]] + theme_bw(base_size=15),filename='BAUviolinspill.png',width=9,height=6.5)
# 
# p[[1]][[1]]

```

```{r violinscen1,fig.cap='Model results for the 100-day SARS-X vaccine without the BPSV.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}
p[[scen_to_keep[2]]]
```

```{r violinscen2,fig.cap='Model results for the 365-day SARS-X vaccine with the BPSV.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}
p[[scen_to_keep[3]]]
```

```{r violinscen3,fig.cap='Model results for the 100-day SARS-X vaccine with the BPSV.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}
p[[scen_to_keep[4]]]
```

```{r violinscen4,fig.cap='Model results for the 100-day SARS-X vaccine with the BPSV.',echo=F,warning=F,message=F,fig.height=5,fig.width=6}
p[[scen_to_keep[5]]]
```






\newpage

# Expected value by policy choice

```{r expectedvaluesbar,fig.cap='Expected values of model results.',echo=F,warning=F,message=F,fig.height=6,fig.width=6,eval=T}

# meltmeantab$scen <- factor((meltmeantab$scen), 
#                                     levels=c('365, No BPSV','365, BPSV',
#                                              '100, No BPSV','100, BPSV'))
setDT(meltmeantab)
# meltmeantab[,cost:=sum(value),by=.(scen,igroup,Policy)]
# meltmeantab[,mincost:=cost==min(cost),by=.(scen,igroup)]
meltmeantab[,mincost:=nmin==max(nmin),by=.(scen,igroup)]
ggplot(subset(meltmeantab,scen%in%scenario_names[scen_to_keep])) + 
      geom_bar(aes(x=factor(Policy,levels=newpolicies,labels=newpolicies),
                   y=value,fill=factor(variable,levels=c('education','yll','gdp'),labels=c('Education','YLL','GDP')),
                   alpha=I(as.numeric(mincost)*.5+.5)),stat='identity') + 
  facet_grid(factor(igroup,levels=c('LLMIC','UMIC','HIC'))~scen) + 
      theme_bw(base_size=15) +
        scale_fill_viridis(discrete=T, name="") +
      labs(x='',y='Loss, % of GDP',fill='Loss type') +
        theme(axis.text.x = element_text(angle = 45,  hjust=1,vjust=1),
              legend.position = 'top') 
    # ggsave(p,filename='figures/expectedcosts.pdf',height=10)
# subset(meltmeantab,igroup=='HIC'&scen=='100, BPSV')

# ggplot(subset(meltmeantab,grepl('Secondary',igroupcountry))) + 
#       geom_bar(aes(x=factor(Policy,levels=c('No Closures','School Closures','Economic Closures','Elimination')),
#                    y=value,fill=factor(variable,levels=c('education','yll','gdp'),labels=c('Education','YLL','GDP')),
#                    alpha=I(as.numeric(mincost)*.5+.5)),stat='identity') + 
#   facet_grid(factor(igroupcountry,labels=c('LLMIC','UMIC','HIC'))~factor(scen,levels=c('BAU','100 days','BPSV'))) + 
#       theme_bw(base_size=15) +
#         scale_fill_viridis(discrete=T, name="",option='inferno',end=.925) +
#       labs(x='',y='Loss, % of GDP',fill='Loss type') +
#         theme(axis.text.x = element_text(angle = 45,  hjust=1,vjust=1),
#               legend.position = 'right') -> p
# 
# ggsave(p,filename = 'expectedvaluesbar.png',height=6,width=8)


```





```{r resultsdiff,echo=F,warning=F,message=F}

difflist <- c()
alldiffs <- list()
for(sl in scen_scens){
  scenario <- scenario_names[sl]
  alldiffs[[length(alldiffs)+1]] <- list()
  for(refsl in bau_scens){
    refscen <- scenario_names[refsl]
    difftab <- as.data.frame(readRDS(paste0('code/results/differences/difftab_',refsl,'_',sl,'.Rds')))
    alldiffs[[length(alldiffs)]][[which(bau_scens==refsl)]] <- difftab
    difftab$Cost = difftab$Cost/difftab$gdp*100;
    difftab$scen <- paste0(refscen,' to ',scenario)
    difftab$fromscen <- refscen
    difftab$toscen <- scenario
    difflist <- rbind(difflist,difftab)
  }
}

```

\newpage

# Violin plot of difference


```{r diffviolin,fig.cap='Model results for losses saved by 100-day and BPS vaccines.',echo=F,warning=F,message=F,fig.height=4,fig.width=6}

vioplot <- reshape2::melt(difflist[,colnames(difflist)%in%c('Cost','igroup','toscen','fromscen')],id.vars=c('igroup','toscen','fromscen'))
vioplot$igroup <- factor(vioplot$igroup,levels=c('LLMIC','UMIC','HIC'))
# vioplot$variable <- factor(vioplot$variable,
#                            levels=c('Cost','YLL','School','GDP_loss'),
#                            labels=c('Total','YLL','Education','GDP'))

ggplot(subset(vioplot,toscen%in%scenario_names[scen_to_keep]&fromscen%in%bau_names),aes(fill=toscen, colour=toscen, y=value, x=toscen)) + 
    geom_violin(position="dodge",scale='width') +
  facet_grid(fromscen~igroup, scales="free") + 
    scale_fill_viridis(discrete=T, name="") +
    scale_colour_viridis(discrete=T, name="") +
    theme_bw(base_size = 15)  +
    xlab("") +
    ylab("Percent of GDP") +
  scale_x_discrete(labels=NULL)+ 
  theme(axis.ticks.x=element_blank(),
        panel.spacing = unit(0.5, "lines"),
        legend.position = 'bottom',
        legend.margin=margin(t=-20))-> p
  # ggsave(p,filename = 'figures/resultsdiff.pdf',width=12,height=10)
  p
  
  

```



\newpage

# Differences

```{r tabresdiff,tab.cap='Value of vaccines: expected difference between BAU and 100 days. Mean values and 95% prediction intervals.',echo=F,warning=F,message=F}

meantab <- setDT(difflist)[,.(yll=mean(YLL),
                        education=mean(School),
                        gdp=mean(GDP_loss)),by=.(igroup,fromscen,toscen)]

difflist[,IQR:=rank(Cost)%in%c(round(nsamples/4):round(3*nsamples/4)),by=.(fromscen,toscen,igroup)]
tabres <- setDT(difflist)[IQR==T,list(costs=paste0(signif(range(Cost),2),collapse='--\u200B')
                        ),by=.(fromscen,toscen,igroup)]
# tabres <- dcast(tabres,formula=igroup+fromscen~toscen,value.var = 'costs',fill = '')

tabres[,igroup:=factor(igroup,levels=c('LLMIC','UMIC','HIC'))]
setorder(tabres,igroup,fromscen)
tabres <- subset(tabres, fromscen!='BAU2')
tabres$fromscen = NULL
colnames(tabres)[1:2] <- c('Scenario','Income group')

# write.csv(tabres,'cepi_results/Delta_LIR_IQR_IL.csv',row.names = F, quote = F)

# subtabres <- subset(tabres,`From scenario`%in%bau_names)[,!colnames(tabres)%in%scenario_names[-c(scen_to_keep)],with=F]

lg <- 'Value of vaccines as a percent of GDP: differences between BAU and scenarios. 75% uncertainty intervals. \\label{tab:tabresdiff}'
# for(i in 3) tabres[[i]] <- gsub('\\.','·',tabres[[i]])

if (!knitr::is_html_output(excludes = "markdown")) { 
  pander(tabres,caption=lg)
}else{
    x <- knitr::kable(tabres,escape=F, "html",caption=lg)
  kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position")
}
```

\newpage

# Best-case scenarios


```{r bestcase,tab.cap='Value of vaccines',echo=F,warning=F,message=F}

plotlist <- readRDS('code/results/exXbars.Rds')
valuetablist <- readRDS('code/results/exXvalues.Rds')
valuetab <- do.call(rbind,valuetablist[[1]])
valuetab2 <- valuetablist[[2]]

colnames(valuetab)[2] = 'Scenario'
countercol = c(rep(rownames(valuetab2)[1:8],each=ncscens))

valuetabcounter <- setDT(subset(valuetab,counter=='counter'))
valuetabcounter$Counterfactual = countercol
rownames(valuetabcounter) <- NULL # rownames(valuetab2)[1:16]
valuetabcounter$`From scenario` <- NULL
valuetabcounter[,lq:=as.numeric(strsplit(Value,'--')[[1]][1]),by=Value]
valuetabcounter[,uq:=as.numeric(strsplit(Value,'--')[[1]][2]),by=Value]
valuetabcounter[,ml:=max(lq),by=Counterfactual]
valuetabcounter[,mu:=max(uq),by=Counterfactual]

vtab = subset(valuetabcounter,ml==lq|mu==uq)[,.(Scenario,Value,Counterfactual)]
bsc <- data.frame(sapply(unique(vtab$Counterfactual),function(x)with(subset(vtab,Counterfactual==x),paste0(Scenario,collapse=', '))))
colnames(bsc) <- 'Best-case scenarios'
bsc$Counterfactual <- rownames(bsc)
rownames(bsc) <- NULL
bsc <- bsc[,2:1]
lg <- paste0('Best-case scenarios: scenarios that have the highest $\\Delta$LIR lower quartile or upper quartile for each counterfactual. \\label{tab:bsc}')
# if (!knitr::is_html_output(excludes = "markdown")) { 
#   cat(pander(bsc,caption=lg))
# }else{
#   x <- knitr::kable(bsc,escape=F, "html",caption=lg)
#   (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
# }

```




\newpage

# Absolute values: LIR and numbers of deaths




```{r absvals,tab.cap='Absolute losses and numbers of deaths of one-in-thirty-years and one-in-one-hundred-years events under the BAU scenario.',echo=F,warning=F,message=F}

valuetabbau1 <- valuetab2[!grepl('BAU2',rownames(valuetab2)),]
rownames(valuetabbau1) <- gsub(', BAU','',rownames(valuetabbau1))

# for(i in 1:2) valuetabbau1[[i]] <- gsub('\\.','·',valuetabbau1[[i]])

lg <- paste0('Absolute losses and numbers of deaths of one-in-X-years events under the BAU scenario. \\label{tab:absvals}')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(valuetabbau1,caption=lg))
}else{
  x <- knitr::kable(valuetabbau1,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}

```



\newpage

# Percent of GDP


```{r ex100gdp,tab.cap='Value of vaccines',echo=F,warning=F,message=F}

valuetabgdp <- setDT(subset(valuetab,counter=='gdp'&`From scenario`!='BAU2'))
valuetabgdp[,counter:=NULL]
valuetabgdp$Counterfactual = sapply(countercol[!grepl('BAU2',countercol)], function(x)gsub(', BAU','',x))
rownames(valuetabgdp) <- NULL # rownames(valuetab2)[1:16]
valuetabgdp$`From scenario` <- NULL
# subvaluetab <- valuetabgdp[,!colnames(valuetabgdp)%in%scenario_names[-c(scen_to_keep)]]
lg <- paste0('Value of vaccine investments in the event of a one-in-X-years pandemic (percent of global GDP). \\label{tab:ex100gdp}')

# for(i in 2) valuetabgdp[[i]] <- gsub('\\.','·',valuetabgdp[[i]])

# if (!knitr::is_html_output(excludes = "markdown")) { 
#   cat(pander(valuetabgdp,caption=lg))
# }else{
#   x <- knitr::kable(valuetabgdp,escape=F, "html",caption=lg)
#   (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
# }

```



\newpage

# Percent of counterfactual


```{r ex100,tab.cap='Value of vaccines',echo=F,warning=F,message=F}
valuetabcounter = subset(valuetabcounter,!grepl('BAU2',Counterfactual))
valuetabcounter$Counterfactual = sapply(countercol[!grepl('BAU2',countercol)], function(x)gsub(', BAU','',x))
valuetabcounter[,counter:=NULL]
valuetabcounter[,lq:=NULL]
valuetabcounter[,uq:=NULL]
valuetabcounter[,ml:=NULL]
valuetabcounter[,mu:=NULL]

# for(i in 2) valuetabcounter[[i]] <- gsub('\\.','·',valuetabcounter[[i]])

valuetabcom = cbind(Return=valuetabcounter$Counterfactual,
                    Scenario=valuetabcounter$Scenario,
                    pccounter=valuetabcounter$Value,
                    pcgdp=valuetabgdp$Value)

# subvaluetab <- valuetabcounter[,!colnames(valuetabcounter)%in%scenario_names[-c(scen_to_keep)]]
lg <- paste0('Value of vaccine investments in the event of a one-in-X-years pandemic (percent of counterfactual). \\label{tab:ex100}')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(valuetabcom,caption=lg))
}else{
  x <- knitr::kable(valuetabcom,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}
```





\newpage

# Distributions of loss

```{r ex30bar,fig.cap='Distributions of values of investments for once-in-thirty-years events.',echo=F,warning=F,message=F,fig.height=8,fig.width=10}
print(plotlist)
```

\newpage

# Expected values


```{r examples,tab.cap='Expected value of investments for different temporal assumptions',echo=F,warning=F,message=F,results='asis'}

examples <- readRDS('code/results/expvals_BAU.Rds')
splitfun = function(str,i)
  as.numeric(strsplit(str,'--\u200B')[[1]][i])
examples[,lq:=splitfun(BAU,1),by=Scenario]
examples[,uq:=splitfun(BAU,2),by=Scenario]

accumulate = function(lq, uq, start_year=2050, end_year=2099, start_gdp=113.8, start_gdp_year=2025, discount=0, growth=0, pv_year=2025){
  years_to_start = start_year - start_gdp_year
  years_to_accumulate = end_year - start_year + 1
  starting_gdp = start_gdp*(1+growth)^years_to_start
  discounted_accumulator = ifelse(discount==growth, years_to_accumulate, (1-(1+discount-growth)^-years_to_accumulate)/(discount-growth))
  discounted_accumulated_gdp = starting_gdp * discounted_accumulator
  pv = discounted_accumulated_gdp/(1+discount)^(start_year-pv_year)
  lqv = lq*pv
  uqv = uq*pv
  ret = paste0(signif(c(lqv, uqv),2),collapse='--\u200B')
  ret
}

examples[,`Example 1`:=accumulate(lq,uq,growth=0.023,discount=0.023),by=Scenario]
examples[,`Example 2`:=accumulate(lq,uq,growth=0.03,discount=0.04),by=Scenario]
examples[,`Example 3`:=accumulate(lq,uq,start_year=2040,growth=0.023,discount=0.023),by=Scenario]
examples[,`Example 4`:=accumulate(lq,uq,start_year=2040,end_year=2089,growth=0.03,discount=0.04),by=Scenario]

examples$BAU <- examples$lq <- examples$uq <- NULL

lg <- paste0('Expected value of investments for different temporal assumptions. \\label{tab:examples}')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(examples,caption=lg))
}else{
  # (knitr::kable(examples,escape=F, "html",caption=lg))
  x <- knitr::kable(examples,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}
```

```{r,eval=F}
subexp <- expvals2#[,!colnames(expvals)%in%scenario_names[-c(scen_to_keep)],with=F]
colnames(subexp)[2] = 'Expected impact'
# for(i in 2) subexp[[i]] <- gsub('\\.','·',subexp[[i]])


lg <- paste0('Value of investments. \\label{tab:expval2}')
if (!knitr::is_html_output(excludes = "markdown")) { 
  cat(pander(subexp,caption=lg))
}else{
  # (knitr::kable(subexp,escape=F, "html",caption=lg))
  x <- knitr::kable(subexp,escape=F, "html",caption=lg)
  (kableExtra::kable_styling(x,full_width = F,latex_options = "HOLD_position"))
}



```


\newpage

# Policy choices

```{r choicestab,fig.cap='Frequency of mitigation policy choices.',echo=F,warning=F,message=F,fig.height=4,fig.width=6}

choicestab <- readRDS(paste0('code/results/choicestab.Rds'))
nsamples <- sum(choicestab$N)/3/nScen
ggplot(subset(choicestab,scenario%in%scenario_names[scen_to_keep])) +  
  geom_bar(aes(x=scenario,
               y=N/nsamples*100,
               fill=factor(policy,levels=strategies,labels=newpolicies)),stat='identity') + 
  facet_wrap(~factor(igroup,levels=c('LLMIC','UMIC','HIC'))) + 
  theme_bw(base_size=15) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_viridis(discrete=T, name="",option='inferno',end=.925,direction=-1) +
  labs(x='',y='Chosen policy, % of samples',fill='') +
  theme(axis.text.x = element_text(angle = 45,  hjust=1,vjust=1),
        legend.position = 'left') 

```


```{r kept,echo=F,warning=F,message=F,eval=T}

setDT(allallresults)
allallresults[,Costrnd:=Cost]
allallresults[Policy!='NC',Costrnd:=Cost*(1+runif(sum(allallresults$Policy!='NC'))*1e-8)]
allallresults[,mincost:=min(Costrnd),by=.(igroup,scenario,samplei)]
# allallresults[,mincost:=min(Cost),by=.(igroup,scenario,samplei)]
allallresults[,keeprow:=Costrnd==mincost]
allallresults$Policy <- factor(allallresults$Policy,levels=newpolicies,labels=newpolicies)

keptresults <- subset(allallresults,keeprow==T)

```


\newpage

## Policy choices vs. severity


```{r intensity,fig.cap='The relationship between policy choice and severity (total deaths).',echo=F,warning=F,message=F,fig.height=5,fig.width=6,eval=T}

keptresults[,qdeath:=rank(Deaths)/.N,by=.(igroup,scenario)]

ggplot(subset(keptresults,scenario%in%scenario_names[scen_to_keep])) + 
  geom_histogram(aes(x=qdeath,y=stat(count / sum(count)),fill=Policy,group=Policy),position='fill',bins = 10) +
  facet_grid(factor(igroup,levels=c('LLMIC','UMIC','HIC'))~scenario) + 
  theme_bw(base_size = 11) +
  scale_x_continuous(expand=c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_viridis_d(name="",option='inferno',end=.925,direction=-1) +
  theme(axis.text.x = element_text(angle = 45,  hjust=1,vjust=1),
        legend.position = 'top') + labs(fill='',x='Deaths per thousand (quantile)',y='Proportion')

```





