---
title: "The Costing Model"
#date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2: 
    #   --filter=pandoc-xnos
    toc: false
    keep_tex: yes
    citation_package: natbib
    extra_dependencies: ["float"]
    # extra_dependencies: ["flafter"]
    # pandoc_args:
    number_sections: true
    fig_caption: yes
    # includes:
    #   in_header: "preamble.tex"
  bookdown::github_document2:
    # pandoc_args: --webtex
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: true
  bookdown::word_document2: 
    toc_depth: 5
    toc_float: true
    number_sections: true
    editor_options: 
      chunk_output_type: inline
bibliography: 
  - "../../epi.bib"
always_allow_html: true
---





<!-- # Figures (temporary) {.unlisted .unnumbered} -->

```{r setup, include=FALSE}
library(ggplot2)  
library(knitr)     
library(tidyr)
library(dplyr)
library(stringi)
library(gplots)
library(RColorBrewer)
library(data.table)
library(splines)
library(bookdown)
library(pander)
library(haven)
library(viridis)
# library(hrbrthemes)
library(MASS)
library(kableExtra)
library(wbstats)
library("cowplot")
library(latex2exp)
library(statmod)
library(TruncExpFam)
library(EnvStats)
library(extraDistr)
library(PearsonDS)
library(ggpubr)

source('functions.R')

panderOptions('round',2)
panderOptions('table.split.table', Inf)

knitr::opts_chunk$set(comment=NA, prompt=FALSE, cache=FALSE, echo=F, message=F, warning=F, results='asis')

format_to_print <- function(x,z=-1){
  formatC(round(x,z), format="f", digits=as.numeric(z>0), big.mark=",")
}

decimalplaces <- function(x) {
    if ((x %% 1) != 0) {
        nchar(strsplit(sub('0+$', '', as.character(format(x,scientific = F))), ".", fixed=TRUE)[[1]][[2]])
    } else {
        return(0)
    }
}

format_to_print2 <- function(x,z=2){
    v <- signif(x,z)
    sapply(v, function(i)  formatC(i, format="f", digits=decimalplaces(i), big.mark=","))
}

```

```{r ages, eval=FALSE}

# age data from owid
allpop = read.csv('population-by-five-year-age-group.csv')
ils = unique(allpop$Entity)[grepl('-income',unique(allpop$Entity))]
maxyear = max(allpop$Year)
ildata = subset(allpop,Entity%in%ils&Year==maxyear)
popcols = which(grepl('Population',colnames(ildata)))
col14 = which(grepl('14',colnames(ildata)))
col64 = which(grepl('64',colnames(ildata)))
for(i in 1:nrow(ildata)){
  print(ildata$Entity[i])
  # print(sum(as.numeric(ildata[i,popcols])))
  print(round(c(sum(as.numeric(ildata[i,popcols[popcols>col14]]))/1e6,
          sum(as.numeric(ildata[i,popcols[popcols>col64]]))/1e6)))
}

```





This document describes the costing model that is used in the CEPI application. It computes the costs of vaccine technology in preparedness and outbreak response, as well as delivery profiles of vaccines by country income level.



# Parameters

```{r paramtable, echo=F,warning=F,message=F}

caps <- readODS::read_ods('cost_parameters.ods',sheet = 1)
caps$`Math notation` <- paste0('$',caps$`Math notation`,'$')
lg <- 'Notation and parametric assumptions for inputs to the costing model. Parameters are used as follows: uniform distributions go from Parameter 1 to Parameter 2. Triangular distributions go from Parameter 1 to Parameter 3 with a peak at Parameter 2. Multinomial distributions have equally probable values listed individually. Exponential distributions have as a mean Parameter 1. Inverse Gaussian distributions have as a mean Parameter 1, and as a shape Parameter 2. Log normal distributions have as a mean Parameter 1, and as a standard deviation Parameter 2. PearsonV distributions have shape Parameter 1, scale Parameter 2, and location 0. PearsonVI distributions have shape Parameters 1 and 2, scale Parameter 3, and location 0.'
capscopy = caps
if (knitr::is_html_output()) {
  capscopy$Source <- gsub('}','',capscopy$Source)
  capscopy$Source <- gsub('\\\\citet\\{','@',capscopy$Source)
} else {
  capscopy$Code <- NULL
  parameter_names = list(Exponential=c('Mean: '),
                         `Inverse Gaussian`=c('Mean: ','Shape: '),
                         `Log normal`= c('Mean: ','Standard deviation: '),
                         PearsonV = c('Shape: ', 'Scale: '),
                         PearsonVI = c('Shape 1: ','Shape 2: ', 'Scale: '))
  for(i in 1:nrow(capscopy)){
    if(capscopy$Distribution[i] %in% names(parameter_names)){
      capscopy$Parameters[i] = paste0(paste0(parameter_names[[capscopy$Distribution[i]]], format_to_print2(as.numeric(strsplit(capscopy$Parameters[i],', ')[[1]]),8)),collapse=', ')
    }
  }
}
cat(pander(capscopy,caption=lg, missing = ""))
i=37
```


```{r sample}

## parameter samples  ######################################################

plist <- get_parameters(nsamples=10000)

params <- plist[[2]]
rawpar <- plist[[1]]

# pardf = as.data.frame(do.call(cbind,rawpar))

```


```{r}

## model variables ######################################################

# weeks_init = sapply(MAN_TYPES,function(x)rawpar[[paste0('weeks_init_',x)]])
# weeks_scale = sapply(MAN_TYPES,function(x)rawpar[[paste0('weeks_scale_',x)]])


# ics = c('hic','umic','lmic','lic')
crs= c(0, 0.7, 2)

## populations and vaccine demand

# pops = sapply(paste0('pop_',ics,'_0'),function(x)rawpar[[x]])
# pops0 = pops[1:3]
# pops0[3] = sum(pops[3:4])
# 
# pops = sapply(paste0('pop_',ics,'_15'),function(x)rawpar[[x]])
# pops3 = pops[1:3]
# pops3[3] = sum(pops[3:4])
# 
# pops65 = sapply(paste0('pop_',ics,'_65'),function(x)rawpar[[x]])
# pops365 = pops65[1:3]
# pops365[3] = sum(pops65[3:4])
# 
# demand65 = rawpar$final_vaccine_coverage*pops365 / (1-rawpar$vaccine_wastage) # total bpsv doses per income level
# popfracs65 = pops365/sum(pops365)
# total_bpsv_demand = 1 # billion. not: sum(demand65/1e9)
# 
# DEMAND15 = rawpar$final_vaccine_coverage*pops3*2 / (1-rawpar$vaccine_wastage)
# popfracs15 = pops3/sum(pops3)
# # two doses plus two boosters
# max_demand = sum(DEMAND15)*2

basic_rates = c(.07, .07, .02, .02)



fifteen_yr_discount_weight = sapply(1:NSAMPLES, function(x) accumulate(discount = pardf$discount[x], from = 1, to = pardf$years_100[x]))
five_yr_discount_weight = sapply(1:NSAMPLES, function(x) accumulate(discount = pardf$discount[x], from = 1, to = pardf$years_200[x]))
at_yr_15 = sapply(1:NSAMPLES, function(x) accumulate(discount = pardf$discount[x], from = pardf$years_100[x]+1, to = pardf$years_100[x]+1))
at_yr_6 = sapply(1:NSAMPLES, function(x) accumulate(discount = pardf$discount[x], from = pardf$years_200[x]+1, to = pardf$years_200[x]+1))

## enabling costs
# matches order of dms
dm_rd_times = c(0, rawpar$years_200, rawpar$years_100)
get_enabling_costs = function(cost_enab, discount=0, target_dm = 365){
  rd_years = dm_rd_times[which(dms==target_dm)]
  cost_enab * accumulate(discount, from=1, to=rd_years)
}

enabling_costs_list = list()
for(x in 1:length(dms)){
  enabling_costs_list[[x]] <- rep(0,NSAMPLES)
  for(i in 1:NSAMPLES){
    enabling_costs_list[[x]][i] = get_enabling_costs(pardf$cost_enab[i], pardf$discount[i], target_dm = dms[x])
  }
}

```


```{r}

## trial timings

get_phase_durations = function(dm=365){
  # phase durations depend on prior R&D investments
  # there are four phases 
  # there are three options for dm: 365, 200 and 100
  phase_duration = rep(0,4)
  for(i in 0:3){
    pars = paste0('weeks_P',i,'_',dm)
    phase_duration[i+1] = rawpar[[pars]]
  }
  phase_duration
}

phase_duration = matrix(0,length(dms),4)
for(j in 1:length(dms)){
  phase_duration[j,] = get_phase_durations(dms[j])
}

# trial durations in normal times
old_durations = do.call(cbind,lapply(paste0('duration_',0:3),function(x)rawpar[[x]]))


timescale = list()
for(j in 1:length(dms)){
  timescale[[j]] = matrix(0,NSAMPLES,4)
  for(i in 1:NSAMPLES) {
    timescale[[j]][i,] = get_duration_scalar(phase_duration[j,],rep(52,4))
  }
}


```

```{r sup}

### bpsv

bpsv_supply = get_bpsv_supply()

b_supply = bpsv_supply$b_supply
b_alloc = bpsv_supply$b_alloc
bpsv_dose_delivery = bpsv_supply$bpsv_dose_delivery
b_tot = max(b_supply$doses)

```


```{r}

# cost per reserved dose upon delivery
cost_res = rawpar$cost_cogs*(1+rawpar$profit)*(1+rawpar$cost_travel)

bpsv_rd_costsamples_dyd <- bpsv_rd_costsamples_no_d <- bpsvresrd <- 
  inv_cost_per_year <- time_to_bpsv <- bpsvproc <- rep(0,NSAMPLES)
bpsv_rd_py = list(); for(i in 1:2) bpsv_rd_py[[i]] = matrix(0,nrow=NSAMPLES,ncol=15)
for(i in 1:NSAMPLES){
  exi = EX[i,]
  inexi = INEX[i,]
  randd_costs = get_bpsv_costs(b_tot, 
                               cost_res = cost_res,
                               pos = POS_BPSV[i,],
                               pto = PTO_BPSV[i,],
                                exi = exi,
                                inexi = inexi,
                                inex_weight = pardf$inex_weight[i],
                                y_durations = old_durations[i,1:3], # years
                                old_duration = old_durations[i,4]/52, # weeks
                                discount = pardf$discount[i],
                                n_bpsv_p1 = 0,#pardf$n_bpsv_p1[i],
                                bpsv_res_upfront = pardf$bpsv_res_upfront[i],
                                n_bpsv_candidates = pardf$n_bpsv_candidates[i],
                                cost_lic = pardf$cost_lic[i],
                                cost_cogs = pardf$cost_cogs[i],
                                cost_ff = pardf$cost_ff[i],
                                cost_travel = pardf$cost_travel[i],
                                profit = pardf$profit[i],
                                bpsv_replenishment = pardf$bpsv_replenishment[i],
                                cost_bpsv_res = pardf$cost_bpsv_res[i])
  
  bpsv_rd_costsamples_dyd[i] = randd_costs[['bpsv_rd_costsamples_dyd']]
  bpsv_rd_costsamples_no_d[i] = randd_costs[['bpsv_rd_costsamples_no_d']]
  bpsv_rd_py[[1]][i,1:length(randd_costs[['bpsv_rd_costsamples_dyd_py']])] = randd_costs[['bpsv_rd_costsamples_dyd_py']]
  bpsv_rd_py[[2]][i,1:length(randd_costs[['bpsv_rd_costsamples_no_d_py']])] = randd_costs[['bpsv_rd_costsamples_no_d_py']]
  bpsvresrd[i] = randd_costs[['bpsvresrd']]
  inv_cost_per_year[i] = randd_costs[['inv_cost_per_year']]
  time_to_bpsv[i] = randd_costs[['time_to_bpsv']]
  bpsvproc[i] = randd_costs[['bpsvproc']]
}

# cost from phase 2 end
##!! wrap
inv_cost = inv_cost_per_year*sapply(1:NSAMPLES,function(x)
  sum(1/(1+pardf$discount[x])^((time_to_bpsv[x]+1):15))
  )
  

```


```{r}

ssv_rd_costsamples <- list()
singlepar <- list()
for(x in 1:length(dms)){
  # ssv_phase_pto[[x]] = matrix(0,NSAMPLES,5)
  ssv_rd_costsamples[[x]] = rep(0,NSAMPLES)
  for(i in 1:NSAMPLES){
    exi = EX[i,]
    cost_lic = pardf$cost_lic[i]
    n_ssv_successes = pardf$n_ssv_successes[i]
    q_ssv_successes = pardf$q_ssv_successes[i]
    randd_costs = get_ssv_randd_costs(timescale = timescale[[x]][i,],
                                      pos = POS_SSV[i,],
                                      pto = PTO_SSV[i,],
                                      exi = exi,
                                      n_ssv_successes = n_ssv_successes,
                                      q_ssv_successes = q_ssv_successes,
                                      cost_lic = cost_lic)
    ssv_rd_costsamples[[x]][i] = randd_costs
  }
}
# summary(ssv_rd_costsamples[[1]])

```

```{r}

# capacity reservation cost

capres_costs_per_year = list()
for(x in 1:length(crs)){
  capres_costs_per_year[[x]] = rep(0,NSAMPLES)
  for(i in 1:NSAMPLES){
    capres_costs_per_year[[x]][i] = get_cap_res_cost(cap_cost_per_dose = pardf$cost_capres[i], 
                                                     cap_res_vol = crs[x])
  }
}


```

```{r}


## scenario variables  ######################################################

scennames = c('BAU',paste0('S',sprintf("%02d", 1:10)))
nscen = length(scennames)
pop_proportional = c(1,1,1,1,1,2,2,1,2,2,2)
scen_dm = c(1, 1, 1,1, 2,2,2, 3,3,3, 1)
scen_capres = c(1, 1, 2,3, 1,2,3, 1,2,3, 1)
bpsv_scen = rep(F,length(scennames))
bpsv_scen[2] = T

rates = matrix(rep(basic_rates, nscen), ,ncol=NLEVELS,byrow=T)
rates[nscen,3:4] = 0.04



```



```{r}
printeq = function(eqtext){
  if (knitr::is_html_output()) {
  cat(paste0("\\begin{equation}\n",gsub("\\\\(?!([a-z]|#))", "\\\\\\\\", eqtext, perl = TRUE),"\n\\end{equation}\n",collapse=''))
} else {
  cat(paste0("\\begin{equation}\n",eqtext,"\\end{equation}\n",collapse=''))
}
}


```

```{r}
printmath = function(eqtext){
  if (knitr::is_html_output()) {
    cat(paste0("```math\n",eqtext,"\n```",collapse=''))
  } else {
    cat(paste0("$$\n",eqtext,"\n$$",collapse=''))
  }
}
```



# Preparedness costs 


We can write the total preparedness cost for scenario $s$ in year $y$, with discount rate $r$, as



```{r}
eqtext = 'D_{s,y}^{\\text{(prep)}} = \\frac{1}{(1+r)^y}\\left(D_s^{\\text{(BP-adRD)}} + D_{s,y}^{\\text{(BP-sec)}} + D_{s,y}^{\\text{(BP-inv)}} + D_s^{\\text{(S-cap)}} + D_{s,y}^{\\text{(en)}}\\right)'
printmath(eqtext)
```

where:

- $D_s^{\text{(BP-adRD)}}$ is the R\&D cost of BPSV prior to an outbreak; see Equation \@ref(eq:bpsvrd)
- $D_{s,y}^{\text{(BP-sec)}}$ is the upfront cost of securing an investigational reserve of `r format_to_print2(rawpar$bpsv_inv_res)` BPSV doses; see Equation \@ref(eq:bpsvadman)
- $D_{s,y}^{\text{(BP-inv)}}$ is the annual cost of maintaining an investigational reserve of `r format_to_print2(rawpar$bpsv_inv_res)` BPSV doses; see Equation \@ref(eq:bpsvinv)
- $D_s^{\text{(S-cap)}}$ is the cost of reserved capacity for SSV; see Equation \@ref(eq:ssvcap)
- $D_{s,y}^{\text{(en)}}$ is the annual cost of enabling activities; see Equation \@ref(eq:enable).

We use $s=0$ to denote the business as usual (BAU) scenario, and $s=1$ for scenario S01,  $s=2$ for scenario S02, and so on.


## BPSV advance R\&D

<!-- **These values match the spreadsheet results** -->


Advance R\&D for BPSVs consists of Phase 0 (which we use to denote preclinical R\&D), Phase 1 and Phase 2, for which we add up costs that depend on (a) the number of candidates, (b) the cost per phase for experienced (those with licensure experience) and inexperienced developers (those without), and (c) the probability of success in each phase. The final cost, therefore, reflects the number of candidates that progressed through the developmental pipeline.

In order to achieve one successful BPSV candidate in the event of an outbreak, we assume that 14 are required to enter into preclinical R\&D [@CEPI2026a]. At present, there are seven (Table \@ref(tab:inex)). We assume that the remaining seven will come from developers without prior licensure experience.

Probabilities of success for Phase 0 and Phase 1 are $P_0^{\text{(BPSV)}}$ and $P_1^{\text{(BPSV)}}$. The probability of phase occurrence for Phase $i$ is:

```{r}
eqtext = '\\hat{P}_i^{(\\text{BPSV})} = \\begin{cases}1 & i=0 \\\\ 
\\prod_{j=0}^{i-1}P_j^{\\text{(BPSV)}} & i\\in\\lbrace 1,2\\rbrace 
\\end{cases}'
printmath(eqtext)

```


<!-- For $N^{\text{(BPSV-1)}} = `r rawpar$n_bpsv_p1`$ candidate(s), which have already been through the preclinical phase, we have -->

<!-- ```{r} -->
<!-- eqtext = '\\hat{P}_i^{(1)} = \\begin{cases}1 & i=1 \\\\  -->
<!-- \\prod_{j=1}^{i-1}P_j^{\\text{(BPSV)}} & i>1  -->
<!-- \\end{cases}' -->
<!-- printmath(eqtext) -->

<!-- ``` -->


The cost of each phase is $T_i$, a weighted average of costs of experienced ($T_i^{(e)}$) and inexperienced ($T_i^{(n)}$) manufacturers. Assuming that $N^{\text{(BPSV)}}=`r rawpar$n_bpsv`$ candidates start in the preclinical phase [@CEPI2026a], and two have prior experience with licensure (Table \@ref(tab:inex)), we take $\omega = `r round(rawpar$inex_weight,2)`$;

<!-- $$T_{i} = (1+\pi)(\omega T_i^{(n)} + (1-\omega)T_i^{(e)}).$$  -->

```{r}
eqtext = 'T_{i} = (1+\\pi)\\left(\\omega T_i^{(n)} + (1-\\omega)T_i^{(e)}\\right)'
printmath(eqtext)
```


where $\pi = `r rawpar$inflation`$ is inflation from 2018 to 2025 [@U.S.BLS2025]. Then the total weighted cost for phases 0 through 2 for $N^{\text{(BPSV)}}$ candidates is



```{r}
eqtext = 'D_s^{\\text{(BP-adRD)}} = \\begin{cases}
 N^{\\text{(BPSV)}}\\sum_{i=0}^2 \\hat{P}_i^{(\\text{BPSV})}T_{i} \\; & \\; s=1 \\\\
0  \\; & \\; s\\neq 1
\\end{cases}
(\\#eq:bpsvrd)'
printeq(eqtext)
```




| Developer | Licensure Experience | 
|---|---|
| Bharat/U Syd/ExcellGene | Yes |
| CPI/CalTech | No | 
| Intravacc | No |
| IVI | No | 
| Panacea/THSTI | No | 
| SK Bio | Yes | 
| VIDO | No | 

Table: (\#tab:inex) Manufacturers working on BPSV and whether or not they have licensure experience as of February 2026 [@CEPI2026a].

 
 
 
```{r posbpsv,fig.cap='Risk-adjusted R&D cost for 14 BPSV candidates'}


ggplot() + geom_histogram(aes(bpsv_rd_costsamples_dyd/1000)) +
  theme_bw(base_size=15) +
  labs(x='Billion USD',y='',title='')


# print(round(summary(bpsv_rd_costsamples_dyd),2))

```

<!-- Target: 333 (232 304 401) -->



## BPSV investigational reserve

```{r}

bpsv_upfront = pardf$bpsv_res_upfront*pfixed$bpsv_inv_res/1e6 # millions


```

<!-- **The annual cost annual is correct, at around 162 thousand, but the total cost is slightly too high** -->


Denoting the duration of Phase $i$ as $Y_i^{(B)}$, the time taken to complete development of the BPSV up to the end of Phase 2, from which point it is manufactured to be held in an investigational reserve, is:


```{r}
eqtext = 'Y^{(B)} = Y_0^{(B)} + Y_1^{(B)} + Y_2^{(B)}.'
printmath(eqtext)
```

The upfront cost of securing the investigational reserve is

```{r}
eqtext = 'D_{s,y}^{\\text{(BP-sec)}} = \\begin{cases}
 A_4A_5
\\; & \\; s=1 \\;\\&\\;y=Y^{(B)}+1\\\\
0  \\; & \\; s\\neq 1\\;\\|\\;y\\neq Y^{(B)}+1
\\end{cases}
(\\#eq:bpsvadman)'
printeq(eqtext)

```

where $A_4 =`r format_to_print(rawpar$bpsv_inv_res)`$ is the size of the reserve and $A_5 =`r format_to_print2(rawpar$bpsv_res_upfront,3)`$ is the cost per dose in USD, coming from the upfront cold-chain equipment cost estimate of 138 million USD for 1.2 billion doses [@Griffiths2021a].

The cost of goods sold (COGS) is $G = `r rawpar$cost_cogs`$ USD per dose. Fill/finish costs from @Kis2021 of 0.29 and 0.39 USD per dose respectively for Moderna and Pfizer doses translate to 7.2% of COGS on average. Then the cost of drug substance, accounting for the fill/finish margin $M_f = `r round(rawpar$cost_ff,2)`$ and the profit margin $M_p = `r round(rawpar$profit,2)`$, is $G(1-M_f)(1+M_p) = `r round(rawpar$cost_cogs*(1-rawpar$cost_ff)*(1+rawpar$profit),2)`$ USD per dose. 

The reserve is replenished every $Y_{rep} = `r rawpar$bpsv_replenishment`$ years based on the storage characteristics for Ervebo [@WorldHealthOrganization2019a]. Then the annual cost to maintain the reserve of $A_4 =`r format_to_print(rawpar$bpsv_inv_res)`$ doses is

```{r}
eqtext = 'D_{s,y}^{\\text{(BP-inv)}} = \\begin{cases}
 \\frac{A_4}{Y_{rep}}G  (1-M_f)(1+M_p) + A_1A_4
\\; & \\; s=1 \\;\\&\\;y>Y^{(B)}\\\\
0  \\; & \\; s\\neq 1\\;\\|\\;y\\leq Y^{(B)}
\\end{cases}
(\\#eq:bpsvinv)'
printeq(eqtext)

```

where $A_1 = `r round(rawpar$cost_bpsv_res,2)`$ USD is the annual reservation cost per dose, coming from the recurring cold-chain cost estimate of 12 million USD for 1.2 billion doses [@Griffiths2021a].

```{r bpsvinv,fig.cap='BPSV investigational reserve costs accumulated from the completion of Phase 2 to year 15 with uniformly distributed discount rate.'}


ggplot() +
  geom_histogram(aes(x=inv_cost)) +
  theme_bw(base_size = 15) +
  labs(x='Cost up to year 15, million USD',y='')


# print(round(summary(inv_cost),2))

```

<!-- Target: 1 (0.9 1 1.1) -->

## SSV capacity reservation

<!-- **This matches the spreadsheet results.** -->


Reservation sizes, in billions, depend on scenarios. It includes the $A_3 = `r rawpar$hic_cap_res`$ billion doses reserved for HIC (325 million for the EU [@Pfizer2023], and 150 million for the USA [@PRNewswire2022], rounded up to 0.5 billion for simplicity of presentation).  The cost of the EU reservation is 160 million EUR per year [@Martuscelli2023]; assuming an exchange rate of 1.08, the cost per dose per year is $A_2 = `r round(rawpar$cost_capres,2)`$ USD. We define the scenario-dependent total reservation size as:

<!-- The cost per dose reservation per year is $A_2 = `r round(rawpar$cost_capres,2)`$ USD. Reservation sizes, in billions, depend on scenarios, including the $A_3 = `r rawpar$hic_cap_res`$ billion doses reserved for HIC, as follows: -->

```{r}
eqtext = 'M_{R,s} = \\begin{cases}A_3 & s\\in\\{0, 1, 4, 7, 10\\} \\\\ 
A_3+0.7 & s\\in\\{2, 5, 8\\} \\\\ 
A_3+2 & s\\in\\{3, 6, 9\\} \\end{cases}'
printmath(eqtext)

```

Then the total cost per year is


```{r}
eqtext = 'D_s^{\\text{(S-cap)}} =  M_{R,s} A_2
(\\#eq:ssvcap)'
printeq(eqtext)

```

The annual costs in billion USD are `r round(0.5*rawpar$cost_capres,2)`, `r round(1.2*rawpar$cost_capres,2)`, and `r round(2.5*rawpar$cost_capres,2)`, respectively.

```{r capres,fig.cap='Capacity reservation costs accumulated over 15 years with uniformly distributed discount rate.'}

##!! wrap
capres_costs = as.data.frame(do.call(rbind,lapply(1:length(crs),function(x) cbind( capres_costs_per_year[[x]]*fifteen_yr_discount_weight, crs[x]))))

ggplot(capres_costs) +
  geom_boxplot(aes(x=factor(V2,levels=crs, labels=crs+.5),y=V1/1e3)) +
  theme_bw(base_size = 15) +
  scale_y_continuous(limits=c(0,NA)) +
  labs(x='Capacity reservation, billion doses',y='Cost, billion USD')


# for(i in unique(capres_costs$V2)){
#   subs = subset(capres_costs,i==V2)
#   cat(i)
#   cat('\n')
#   print(round(summary(subs$V1/1e3),2))
#   cat('\n')
# }

```

<!-- Targets: -->
<!-- 3,086	(2,897 3,074 3,269) -->

<!-- 7,407	(6,954 7,378 7,845) -->

<!-- 15,431	(14,487	15,370	16,344) -->
		



## Enabling activities

<!-- **This matches the spreadsheet results** -->


Denote the target "days to SSV" by $\zeta$, so that $\zeta\in\lbrace 365, 200, 100 \rbrace$, which are the three variations we consider enabling activities might achieve. We assume that $Y^{(200)}=`r rawpar$years_200`$ years are required to achieve 200 days to SSV, and $Y^{(100)}=`r rawpar$years_100`$ years are required to achieve 100 days to SSV. Then annual costs, $E=700$ million USD, accumulate depending on the year and the target:

```{r}
eqtext = 'D_{s,y}^{\\text{(en)}} = \\begin{cases}E & \\zeta(s)=200 \\;\\&\\; y\\leq Y^{(200)} \\; |\\; \\zeta(s)=100\\; \\& \\;y\\leq Y^{(100)} \\\\ 
0 & \\zeta(s)=365 \\;|\\; y > Y^{(100)} \\;|\\; \\zeta(s)=200 \\;\\&\\; y \\;>\\; Y^{(200)}  \\end{cases}
(\\#eq:enable)'
printeq(eqtext)

```

We use $\zeta(s)$ to denote the mapping from our scenario definitions to their respective $\zeta$ values as follows:


```{r}
eqtext = '\\zeta(s) = \\begin{cases} 365 & s\\in\\{0, 1, 2, 3, 10\\} \\\\ 
200 & s\\in\\{4, 5, 6\\} \\\\ 
100 & s\\in\\{7, 8, 9\\} \\end{cases}'
printmath(eqtext)

```


```{r en,fig.cap='Enabling costs accumulated over 15 years with uniformly distributed discount rate.'}


enabling_costs = as.data.frame(do.call(rbind,lapply(1:length(dms),function(x) cbind(dms[x],  enabling_costs_list[[x]]))))

ggplot(enabling_costs) +
  geom_boxplot(aes(x=factor(V1,levels=dms),y=V2/1e3)) +
  theme_bw(base_size = 15) +
  labs(x='Days Mission',y='Cost, billion USD')
# 
# for(i in unique(enabling_costs$V1)){
#   subs = subset(enabling_costs,i==V1)
#   cat(i)
#   cat('\n')
#   print(round(summary(subs$V2/1e3),2))
#   cat('\n')
# }


```

<!-- Targets: -->

<!-- 3,242	(3,182 3,241 3,302) -->

<!-- 8,126 (7,629 8,094 8,607) -->


# Response cost equation 


We write total response costs as



```{r}
eqtext = 'D_{s,y}^{\\text{(res)}} = \\frac{1}{(1+r)^y}\\left(D_{s}^{\\text{(S-RD)}} + D_{s}^{\\text{(BP-resRD)}} + D_{s,y}^{\\text{(S-proc)}} + D_{s}^{\\text{(BP-proc)}} + D_{s,y}^{\\text{(S-del)}} + D_{s,y}^{\\text{(BP-del)}}\\right)'
printmath(eqtext)
```

where

- $D_{s}^{\text{(S-RD)}}$ is the R\&D cost for SSV; see Equation \@ref(eq:ssvrd)
- $D_{s}^{\text{(BP-resRD)}}$ is the R\&D cost of BPSV after an outbreak; see Equation \@ref(eq:bpsvresrd)
- $D_{s,y}^{\text{(S-proc)}}$ is the cost of procuring SSV; see Equation \@ref(eq:ssvproc)
- $D_{s}^{\text{(BP-proc)}}$ is the cost of procuring BPSV; see Equation \@ref(eq:bpsvproc)
- $D_{s,y}^{\text{(S-del)}}$ is the cost of delivering SSV; see Equation \@ref(eq:ssvdel)
- $D_{s,y}^{\text{(BP-del)}}$ is the cost of delivering BPSV; see Equation \@ref(eq:bspvdel)

We present computation for this cost with reference to the beginning of the outbreak. Therefore, indices $y$ in this section do not correspond to the same index in the previous section, where $y$ indexed years of investment pre-pandemic.



## Risk-adjusted R&D cost per candidate calculation


### SSV

<!-- **These don't match the spreadsheet results. Values too low.** -->

Trial costs are adjusted for the duration of the trial, which depend on the R\&D enabling investment, denoted $\zeta\in\lbrace 365, 200, 100\rbrace$: 



```{r}
eqtext = 'T_{\\zeta,i}^{(e)} = (1+\\pi)\\frac{W_i^{(\\zeta)}}{52Y_{i}^{(B)}}T_i^{(e)}'
printmath(eqtext)
```


where $\pi = `r rawpar$inflation`$ is inflation from 2018 to 2025, $T_i^{(e)}$ is the cost per Phase $i$ of experienced developers (i.e. we assume only experienced manufacturers undertake reactive R\&D), $Y_{i}^{(B)}$ is the usual phase duration in normal times in years, and $W_i^{(\zeta)}$ is its expected duration in weeks given enabling investments made prior to the outbreak. 

The probability of success of each phase comes from COVID-19 data [@LinksbridgeSPC2025]:


```{r}
eqtext = 'P_i^{\\text{(SSV)}} \\sim \\text{Beta}\\left(\\sum_{j=i+1}^4X_j+1, X_i + 1 \\right)'
printmath(eqtext)

```

for $i \in \lbrace 0,1,2,3 \rbrace$ where $X_i$ is the number of candidates that failed in phase $i$ and $X_4$ the number that succeeded to licensure.

The probabilities of phase occurrence are:

```{r}
eqtext = '\\hat{P}_i^{\\text{(SSV)}} = \\begin{cases}1 & i=0 \\\\ 
\\prod_{j=0}^{i-1}P_j^{\\text{(SSV)}} & i\\in\\lbrace 1,2,3,4\\rbrace 
\\end{cases}'
printmath(eqtext)

```



Then the total cost is


```{r}
eqtext = 'D_s^{\\text{(S-RD)}} = N^{\\text{(SSV)}}\\left(\\sum_{i=0}^3 \\hat{P}_i^{\\text{(SSV)}} \\cdot T_{\\zeta(s),i}^{(e)} +  \\hat{P}_4^{\\text{(SSV)}} T_4\\right)
(\\#eq:ssvrd)'
printeq(eqtext)

```



where $T_4$ is the cost of licensure. We multiply by the number of candidates, $N^{\text{(SSV)}}$, to get the total cost from the weighted average per candidate, where


```{r}
eqtext = 'N^{\\text{(SSV)}} = n^{\\text{(SSV)}} + F^{-1}_{NegBin}\\left(Q^{\\text{(SSV)}}; n^{\\text{(SSV)}},  \\hat{P}_4^{\\text{(SSV)}} \\right)'
printmath(eqtext)

```

is chosen to secure at least $n^{\text{(SSV)}} = `r rawpar$n_ssv_successes`$ successful candidates with probability $Q^{\text{(SSV)}} = `r 100*rawpar$q_ssv_successes`$%. Here, $F^{-1}_{NegBin}\left(q; n,  p \right)$ is the cumulative density of a negative binomial distribution with parameters $n$ and $p$ evaluated at quantile $q$.

```{r posssv,fig.cap='Risk-adjusted R&D cost to secure at least five successful SSV candidates'}

# prep for plot
ssv_rd_costsamples_lab = lapply( 1:length(ssv_rd_costsamples), function(x) cbind( ssv_rd_costsamples[[x]], paste0(dms[x], ' Days Mission')) )

ssv_rd_plot <- as.data.frame(do.call(rbind,ssv_rd_costsamples_lab))
colnames(ssv_rd_plot) = c('Cost','Investment')
ssv_rd_plot$Cost = as.numeric(ssv_rd_plot$Cost)
ggplot(ssv_rd_plot) + geom_histogram(aes(Cost/1e3)) +
  facet_wrap(~Investment) +
  theme_bw(base_size=15) +
  labs(x='Billion USD',y='',title='')


# pander(as.data.frame(cbind(DM=dms,do.call(rbind,lapply(unique(ssv_rd_plot$Investment), function(i){
#   subs = subset(ssv_rd_plot, Investment==i)
#   round(summary(subs$Cost/(1+pardf$discount)^15))
# }
# )))))

```


<!-- Targets: -->

<!-- 250 (107 170 275) -->

<!-- 153 (54 83 135) -->

<!-- 85 (27 49 88) -->




### BPSV

<!-- **This is a little higher than the spreadsheet results** -->




One BPSV candidate that has passed through phases 0 to 2 prior to the outbreak goes through Phase 3 during the response. The duration is $W_3^{(365)}=`r rawpar$weeks_P3_365`$ weeks. Thus we write the BPSV R\&D response cost


```{r}
eqtext = 'D_s^{\\text{(BP-resRD)}} = \\begin{cases}\\left( (1+\\pi)\\frac{W_3^{(365)}}{52Y_3^{(B)}}T_3^{(e)} + T_4\\right) \\; & \\; s=1 \\\\
0  \\; & \\; s\\neq 1
\\end{cases}
(\\#eq:bpsvresrd)'
printeq(eqtext)

```


```{r bpsvresrd,fig.cap='Reactive R&D cost for BPSV'}

ggplot() + geom_histogram(aes(bpsvresrd)) +
  theme_bw(base_size=15) +
  labs(x='Cost, million USD',y='',title='')

# print(round(summary(bpsvresrd*at_yr_15),1))

```

<!-- Target: 11 (2 4 8) -->



## Procurement cost calculation

The cost per dose comes from the cost of goods sold (COGS), $G = `r rawpar$cost_cogs`$ USD, adjusted for a profit margin ($M_p = `r rawpar$profit`$) and transportation margin ($M_t = `r rawpar$cost_travel`$). The 20% profit margin is at the top of range in @U.S.GovernmentAccountabilityOffice2017, and thus acts as a conservative estimate on profit margins sought by pharmaceutical companies. The 10% estimate for transportation from manufacturer to country isolates the transportation cost from the process of vaccination [@Gotham2019].

$S_R = G(1+M_p)(1+M_t)$ evaluates to `r round(cost_res,2)` USD.

This cost is used both for SSV doses procured via reserved capacity, and all newly manufactured BPSV doses.

### SSV

<!-- **These values are close, but not identical, to the spreadsheet results if I adjust for the total demand** -->

We denote $A_{x,s,y}$ to be the number of doses (in billions) produced from channel $x$ in year $y$ and scenario $s$, where $x\in\lbrace R,E,B \rbrace$, corresponding to reserved, existing (and unreserved), and reactively built capacity, respectively.  $A_{x,s,y}$ comes from Equation \@ref(eq:supply). Then the total cost, in billion USD, is:


```{r}
eqtext = 'D_{s,y}^{\\text{(S-proc)}} = A_{R,s,y} S_R  + \\sum_{x\\in\\lbrace E,B \\rbrace}  A_{x,s,y} S_U
(\\#eq:ssvproc)'
printeq(eqtext)

```

Here, $S_R = `r round(cost_res,2)`$ is the cost per reserved dose and $S_U = `r round(rawpar$cost_un,2)`$ the cost per unreserved dose in USD. 




```{r ssvsup}

supplies <- cumulative <- list()

for(s in 1:nscen){
  ssv_supply = get_ssv_supply(dm=dms[scen_dm[s]], 
                              capres=crs[scen_capres[s]], 
                              bpsv=bpsv_scen[s],
                              weeks_init = WEEKS_INIT,
                              weeks_scale = WEEKS_SCALE)
  
  supplies[[s]] = ssv_supply[['supplies']]
  cumulative[[s]] = ssv_supply[['cumulative']]
}


```

```{r}

# some variables need to be fixed atm as they feed into delivery trajectories
pfixed = list()
pfixed$hic_cap_res = rawpar$hic_cap_res

cost_per_year <- ssv_proccost_discounted <- ssv_proccost_undiscounted <- list()

## ssv allocation 

function_list = list(res=allocate_res_doses, ex=allocate_exbui_doses, bui=allocate_exbui_doses)
rep_res = list(res=allocate_res_doses, ex=allocate_res_doses, bui=allocate_res_doses)

time_to_approval = floor(dms/7)
alloc <- second_dose_delivery <- all_dose_delivery <- list()


for(s in 1:nscen){
  cost_per_year[[s]] = matrix(0, nrow=NSAMPLES, ncol=NYEARS+1)
  ssv_proccost_discounted[[s]] <- ssv_proccost_undiscounted[[s]] <- rep(0, NSAMPLES)
  
  # allocation
  allocation_functions = list(function_list, rep_res)[[pop_proportional[s]]]

  alloc_and_del = allocate_and_deliver_doses(allocation_functions, supplies[[s]], del_rates=rates[s,],
                                             time_to_approval=time_to_approval[scen_dm[s]], 
                                           hic_only = pfixed$hic_cap_res/(crs[scen_capres[s]]+pfixed$hic_cap_res))
  
  alloc[[s]] = alloc_and_del[['alloc']]
  alloc[[s]]$Scenario = scennames[s]
  second_dose_delivery[[s]] = alloc_and_del[['second_dose_delivery']]
  all_dose_delivery[[s]] = alloc_and_del[['all_dose_delivery']]
  
  
  total_doses = annualise_ssv_procurement(supplies[[s]], alloc[[s]], cap_res=pfixed$hic_cap_res + crs[scen_capres[s]])
  for(i in 1:NSAMPLES){
    cost_un = pardf$cost_un[i]
    cost_res = pardf$cost_cogs[i]*(1+pardf$profit[i])*(1+pardf$cost_travel[i])
    ssv_procurement_costs = get_ssv_procurement_costs(ssv_py_inc_booster=total_doses, 
                                                      cost_res=cost_res, 
                                                      cost_un=cost_un)
    
    
    ssv_proccost_discounted[[s]][i] = sum(sapply(1:NYEARS,function(x) ssv_procurement_costs/(1+pardf$discount[i])^x))
    ssv_proccost_undiscounted[[s]][i] = sum(ssv_procurement_costs)/1e3 # billions to millions
  }
  cost_per_year[[s]] = c(ssv_procurement_costs, scennames[s])
}

```


```{r costperyear,fig.cap='SSV procurement cost'}


cpy = data.frame(do.call(rbind,cost_per_year))
colnames(cpy) = c(paste0('Year ',1:NYEARS),'Scenario')
cpymelt = reshape2::melt(cpy,id.vars='Scenario')
cpymelt$variable = as.numeric(cpymelt$variable)
cpymelt$value = as.numeric(cpymelt$value)

ggplot(cpymelt) + 
  geom_line(aes(x=variable,y=value/1e3,colour=Scenario)) +
  theme_bw(base_size=15) +
  labs(x='Year',y='Procurement cost, billion USD')


# pander(as.data.frame(
#   cbind(Scenario=scennames,
#         do.call(rbind,
#                 lapply(cost_per_year, function(x)
#                   round(summary(rowSums(
#                     sapply(1:NYEARS,function(y)
#                       as.numeric(x[y])/(1+rawpar$discount)^c(14+y)))
#                     ))
#                   )
#                 ))),
#   caption = 'Costs summed and discounted from year 16 to year 20, million USD')

```

<!-- Targets: -->

<!--  BAU 	$213,204	$177,135	$209,062	$247,161 -->

<!-- S01	$216,458	$180,402	$212,386	$250,435 -->

<!-- S02	$195,822	$162,791	$192,045	$226,936 -->

<!-- S03	$157,183	$130,875	$154,197	$181,966 -->

<!-- S04	$218,874	$183,133	$214,909	$252,596 -->

<!-- S05	$198,810	$166,431	$195,224	$229,363 -->

<!-- S06	$169,394	$141,909	$166,360	$195,333 -->

<!-- S07	$220,033	$184,390	$216,099	$253,673 -->

<!-- S08	$197,578	$165,656	$194,070	$227,711 -->

<!-- S09	$165,769	$139,125	$162,855	$190,932 -->

<!-- S10	$207,642	$172,542	$203,619	$240,702 -->






### BPSV

<!-- **This is pretty close** -->

The cost of BPSV doses is the sum of new doses supplied, $A_{BPSV,s}$, at the reserved-capacity cost per dose, and the doses held in the investigational reserve, for which fill/finish, transport and profit margin costs are due.


```{r}
eqtext = 'D_{s}^{\\text{(BP-proc)}} = \\begin{cases}
A_{BPSV,s}\\cdot S_R +  A_4(M_f+M_t)(1+M_p)G\\; & \\; s=1 \\\\
0  \\; & \\; s\\neq 1
\\end{cases}
(\\#eq:bpsvproc)'
printeq(eqtext)

```


For a world population aged 65 and over of `r round(sum(POPS65)/1e9,2)` billion, a coverage of `r 100*rawpar$final_vaccine_coverage`\% (accounting for wastage of `r round(100*rawpar$vaccine_wastage)`\%), and a cost per dose of $S_R = `r round(cost_res,2)`$ USD (the same as for SSV via reserved capacity), the procurement cost for BPSV is `r round(bpsvproc[1]/1e3,2)` billion USD.


```{r}
# print(round(summary(bpsvproc*at_yr_15),2))
```

<!-- Target: 3,130 (2,671 3,084 3,565) -->


## Delivery Cost Equation


### SSV

<!-- **These values are ballpark correct but too concentrated**. corrected via perfect correlation between costs. -->

For populations aged 15 and above, $N_i^{(15)}$ for income group $i$, we write 


```{r}
eqtext = 'L_i = 2\\cdot \\lambda\\cdot N_i^{(15)} / (1-\\delta)'
printmath(eqtext)
```

as the total demand for first-schedule doses for income group $i\in\lbrace\text{LIC, LMIC, UMIC, HIC}\rbrace$, representing two doses each for $\lambda = `r rawpar$final_vaccine_coverage*100`$\% of the population and $\delta = `r round(rawpar$vaccine_wastage*100)`$\% wastage.

We write the delivery cost $H_{s,i,w}$ for $h_{s,i,w}$ doses given in week $w$ and income group $i$ as follows. There are three cost tiers, the first of which, $V_{i; 0}$, is applied to the first 10\% of $L_i$ (the "start up" cost); the second ($V_{i; 11}$) to the subsequent 20\% (the "ramp up" cost); and the third ($V_{i; 31}$) to all doses thereafter (the "at scale" cost). The same costing schedule applies both to the first-schedule plus booster SSV doses and the BPSV rollout. 

```{r}
eqtext = 'H_{s,i,w} = 
\\begin{cases}
V_{i; 0}h_{s,i,w}  & \\sum_{j=1}^{w-1}h_{s,i,w}\\leq \\frac{1}{10}L_i \\\\
V_{i; 10}h_{s,i,w}  & \\frac{1}{10}L_i < \\sum_{j=1}^{w-1}h_{s,i,w}\\leq \\frac{3}{10}L_i  \\\\
V_{i; 30}h_{s,i,w}  & \\frac{3}{10}L_i < \\sum_{j=1}^{w-1}h_{s,i,w}
\\end{cases}'
printmath(eqtext)

```


Then the delivery cost in year $y$ and scenario $s$ is

```{r}
eqtext = 'D_{s,y}^{\\text{(S-del)}} = 
\\sum_{w\\in y} \\sum_i H_{s,i,w}
(\\#eq:ssvdel)'
printeq(eqtext)

```




```{r ssvallocation}


# 
# ggplot(subset(do.call(rbind,second_dose_delivery),Scenario%in%scennames)) +
#   geom_line(aes(x=Week,colour=Scenario,y=doses)) +
#   facet_wrap(~factor(il,levels=INCOMELEVELS)) +
#   theme_bw(base_size = 15) +
#   scale_x_continuous(limits=c(0,100)) +
#   labs(x='Week',y='Second-dose coverage, percent population',colour='Scenario')

```


```{r deliverycost,fig.cap='SSV delivery cost'}

## cost delivery

# apply costs per dose to quantiles
##!! can just apply values to the total (rather than going week by week) if there is no discounting
ssv_delivery_costs_list = list()
for(s in 1:length(scennames)){
  ssv_delivery_costs_list[[s]] = rep(0,NSAMPLES)
  for(i in 1:NSAMPLES){
    # cost_per_dose_mat = matrix(0,nrow=length(INCOMELEVELS), ncol=length(PC_THRESHOLDS))
    # for(il in 1:length(INCOMELEVELS))
    #   for(j in 1:length(PC_THRESHOLDS)){
    #     lab = paste0('cost_',tolower(INCOMELEVELS)[il],'_',as.numeric(PC_THRESHOLDS[j]))
    #     cost_per_dose_mat[il,j] = cost_per_dose[[lab]][i]
    #   }
        
    cost_per_dose = COST_PER_DOSE_LIST[[i]]
    discount = pardf$discount[i]
    
    ssv_delivery_costs = get_ssv_delivery_costs(all_ssv_delivered = all_dose_delivery[[s]], 
                                                     cost_per_dose = cost_per_dose,
                                                     discount = discount)
    dis_ssv_delivery_costs = ssv_delivery_costs[['dis_countrycosts']]
    # ssv_delivery_costs = ssv_delivery_costs[['countrycosts']]
    
    ssv_delivery_costs_list[[s]][i] = dis_ssv_delivery_costs
  }
}

ssv_delivery_costs = as.data.frame(do.call(cbind,ssv_delivery_costs_list))
colnames(ssv_delivery_costs) = scennames

ggplot(reshape2::melt(ssv_delivery_costs)) +
  geom_boxplot(aes(x=variable,y=value/1e3,colour=variable),show.legend=F) +
  theme_bw(base_size=15) +
  labs(x='',y='SSV delivery cost, billion USD')

# print(t(summary(ssv_delivery_costs*at_yr_15)))

```

<!-- Targets: -->

<!--  BAU 	$115,985	$92,999	$112,301	$134,660 -->

<!-- S01	$116,404	$93,210	$112,431	$135,476 -->

<!-- S02	$115,989	$93,194	$112,640	$134,450 -->

<!-- S03	$116,234	$93,153	$112,264	$135,306 -->

<!-- S04	$116,699	$94,120	$113,068	$134,927 -->

<!-- S05	$117,038	$94,169	$112,895	$135,891 -->

<!-- S06	$117,222	$94,111	$113,502	$136,028 -->

<!-- S07	$118,839	$95,667	$115,147	$137,742 -->

<!-- S08	$117,980	$94,897	$113,822	$136,699 -->

<!-- S09	$118,170	$95,625	$114,272	$136,928 -->

<!-- S10	$114,979	$92,120	$110,950	$133,834 -->



### BPSV

<!-- **These values match the spreadsheet results. (NB: more doses are purchased and delivered than there are eligible people in the population)** -->


For the BPSV, which goes only to people aged 65 or older, with populations $N_i^{(65)}$, coverage is reached earlier in the process, so the cost is weighted more heavily towards start up and ramp up:


```{r}
eqtext = 'D_s^{\\text{(BP-del)}} = 
\\begin{cases}
\\sum_{i}D_{\\text{BPSV},i}
\\; & \\; s=1 \\\\
0  \\; & \\; s\\neq 1
\\end{cases}
(\\#eq:bspvdel)'
printeq(eqtext)
```



```{r}
eqtext = 'D_{\\text{BPSV},i} = 
\\begin{cases}
N_i^{(65)}V_{i; 0}  & N_i^{(65)}\\leq \\frac{1}{10}N_i^{(15)} \\\\
\\frac{N_i^{(15)}}{10} V_{i; 0} + \\left(N_i^{(65)}-\\frac{N_i^{(15)}}{10} \\right)V_{i; 11}  & \\frac{1}{10}N_i^{(15)} < N_i^{(65)}\\leq \\frac{3}{10}N_i^{(15)} \\\\
\\frac{N_i^{(15)}}{10} V_{i; 0} + \\frac{2}{10}N_i^{(15)} V_{i; 11} + \\left(N_i^{(65)}-\\frac{3}{10}N_i^{(15)} \\right)V_{i; 31} & N_i^{(65)}> \\frac{3}{10} N_i^{(15)}
\\end{cases}'
printmath(eqtext)
```

The logic of this is as follows:

- The increments in cost correspond to numbers of eligible people in the whole population, namely those aged 15 and above.
- If the number of people eligible for the BPSV is less than 10\% of the population aged 15 and over, then all doses cost the "start up" amount.
- If the number of people eligible for the BPSV is more than 10\% and less than 30\% of the 15+ population, then the cost of the first doses, a number equal to 10\% of the 15+ population, is the "start up" amount. All remaining doses cost the "ramp up" amount.
- If the number of people eligible for the BPSV is more than 30\% of the 15+ population, then the cost of the first doses, a number equal to 10\% of the 15+ population, is the "start up" amount. The cost of the second tranche of doses, a number equal to 20\% of the 15+ population, is the "ramp up" amount. All remaining doses cost the "at scale" amount.


```{r bpsvdeliverycost,fig.cap="BPSV delivery cost"}

# points at which costs shift per income level (ignoring the 80% mark)
# del_cost_thresholds = sapply(c(1,3)/10,function(x) x*pops3) 

bpsv_del_cost = rep(0,NSAMPLES)
for(i in 1:NSAMPLES){
  cost_per_dose_mat = COST_PER_DOSE_LIST[[i]]
  bpsv_del_cost[i] = get_bpsv_del_costs(bpsv_dose_delivery=bpsv_dose_delivery, 
                                        cost_per_dose_mat=cost_per_dose_mat)
}

ggplot() + geom_histogram(aes(x=bpsv_del_cost/1e3)) +
  theme_bw(base_size = 15) +
  labs(x='BPSV delivery cost, billion USD',y='')

# print(round(summary(bpsv_del_cost*at_yr_15),2))
  
```


<!-- Target: 9,873 (7,999 9,563 11,447) -->



| Country | Country status | Study type | Financial cost per dose (USD) | Source |
|---|---|---|---|---|
| WHO, Gavi, and UNICEF AMC Estimate | AMC | Top down | 1.66 | @Griffiths2021a |
| UNICEF Global Estimate |  All | Model | 0.73 | @Oyatoye2023 |
| DRC | LIC | Bottom up | 1.91 | @Moi2024 |
| Malawi | LIC | Bottom up | 4.55 | @Ruisch2025 |
| Mozambique | LIC | Bottom up | 0.5 | @Namalela2025 |
| Uganda | LIC | Bottom up | 0.79 | @Tumusiime2024 |
| Bangladesh | LMIC | Bottom up | 0.29 | @Yesmin2024 |
| Cote d'Ivoire | LMIC | Bottom up | 0.67 | @Vaughan2023 |
| Nigeria | LMIC | Bottom up | 0.84 | @Noh2024 |
| Philippines | LMIC | Bottom up | 2.16 | @Banks2023 |
| Vietnam | LMIC | Bottom up | 1.73 | @Nguyen2024 |
| Ghana | LMIC | CVIC tool | 2.2--2.3 | @Nonvignon2022 |
| Lao PDR | LMIC | CVIC tool | 0.79--0.81 | @Yeung2023 |
| Kenya | LMIC | Top down | 3.29--4.28 | @Orangi2022 | 
| Botswana | UMIC | Mixed | 19 | @Vaughan2025 |
| South Africa | UMIC | Top down | 3.84 | @Edoka2024 | 

Table: (\#tab:delcosts) Literature review of global and country-specific delivery costs 






```{r}

missions = paste0(dms[2:3],' days to SSV')
capas = paste0(crs[2:3],' billion capacity')

```




```{r}

## bringing together preparedness costs

iqrs = c(1:3)/4

# bpsv_rd_costsamples_dyd
bpsv_rd_prep = as.data.frame(t(quantile(bpsv_rd_costsamples_no_d/1e3,iqrs)))
bpsv_rd_prep$timing = 'Upfront'
bpsv_rd_prep$category = 'R&D'
bpsv_rd_prep$type = 'BPSV'


# enabling_costs
enab = as.data.frame(t(sapply(c(5,15),function(x) quantile(rawpar$cost_enab*x,iqrs)/1e3)))
enab$timing = 'Upfront'
enab$category = 'R&D'
enab$type = missions

# rbind(enab, bpsv_rd_prep)

# upfront cost in million dollars
bpsv_man = as.data.frame(t(quantile(bpsv_upfront/1e3,iqrs)))
bpsv_man$timing = 'Upfront'
bpsv_man$category = 'Procurement'
bpsv_man$type = 'BPSV'

# inv_cost in millions
bpsv_inv_prep = as.data.frame(t(quantile(inv_cost_per_year,iqrs)))
bpsv_inv_prep$timing = 'Per year'
bpsv_inv_prep$category = 'Procurement'
bpsv_inv_prep$type = 'BPSV'


# capres_costs
ssv_capres = as.data.frame(t(sapply(crs[2:3],function(x) quantile(x*rawpar$cost_capres,iqrs))))
ssv_capres$timing = 'Per year'
ssv_capres$category = 'Procurement'
ssv_capres$type = capas

timings = c('Upfront','Per year','Per pandemic')
types = c('BPSV', missions, capas, 'Equality + Delivery')
prepdata = rbind(enab, bpsv_rd_prep,bpsv_man,bpsv_inv_prep, ssv_capres)
prepdata$timing = factor(prepdata$timing,levels=timings,labels=timings)
prepdata$type = factor(prepdata$type,levels=types,labels=types)
# ggplot(prepdata) + theme_bw(base_size = 15) +
#   geom_hline(yintercept=0,colour='grey') +
#   geom_vline(xintercept=0,colour='grey') +
#   facet_wrap(~timing, scales='free') +
#   geom_bar(aes(x=type,y=`50%`,fill=category),stat = 'identity', position = 'dodge') +
#   geom_errorbar(position = position_dodge(width = 0.9), width=0.6,aes(x=type,group=category,ymin=`25%`,ymax=`75%`),show.legend = F) +
#       theme(axis.text.x = element_text(angle = 60,  hjust=1,vjust=1)) +
#   labs(x='',y='Million USD',fill='',size='',title='Costs vs. BAU')

```


```{r}
# enabling r&d accumulating over time

costperyear = rawpar$cost_enab/1e3
d100 = rep(costperyear,15)
d200 = c(rep(costperyear,5),rep(0,10))


years = 1:15
x=1
d100d = t(sapply(years,function(x) quantile(d100[x]/(1+pardf$discount)^(x-1),c(1,3)/4)))
d200d = t(sapply(years,function(x) quantile(d200[x]/(1+pardf$discount)^(x-1),c(1,3)/4)))

df = (rbind(data.frame(years, m=cumsum(d100), apply(d100d,2,cumsum), d='100'), 
      data.frame(years, m=cumsum(d200), apply(d200d,2,cumsum), d='200'),
      data.frame(years, m=0, X25. = 0, X75. = 0, d='365')))

rdplot = ggplot(df)  +
  geom_ribbon(aes(x=years-1,ymin=(X25.),ymax=(X75.), fill=d, group=d),alpha=0.5, colour=NA,show.legend = F) + 
  geom_line(aes(x=years-1,y=(m),colour=d)) + 
  scale_x_continuous(n.breaks = 8)+# = seq(0,to=max(years),by=3)) +
  theme_bw(base_size = 16) + 
  labs(x='Year',y='Billion USD',colour='Days to SSV') +
  theme(legend.position = 'top')
ggsave(rdplot, filename='../../cepi_results/rdplot.png')


costperyear = sapply(crs+pfixed$hic_cap_res, function(x) rawpar$cost_capres*x)
df = do.call(rbind, lapply(1:length(crs),function(y)
  data.frame(
    apply(t(sapply(years,function(x) quantile(costperyear[y]/(1+pardf$discount)^(x-1),c(1,3)/4))),2,cumsum), 
    m=cumsum(rep(costperyear[y],15)),years, d=as.character(crs[y]+pfixed$hic_cap_res))
))

capresplot = ggplot(df)  +
  geom_ribbon(aes(x=years-1,ymin=(X25.),ymax=(X75.), fill=d, group=d),alpha=0.5, colour=NA,show.legend = F) + 
  geom_line(aes(x=years-1,y=(m),colour=d,group=d)) +
  scale_x_continuous(n.breaks = 8)+# = seq(0,to=max(years),by=3)) +
  theme_bw(base_size = 16) + 
  labs(x='Year',y='Billion USD',colour='Total reserved capacity, billion doses') +
  theme(legend.position = 'top')
ggsave(capresplot, filename='../../cepi_results/capresplot.png')



# bpsv. first, discounted. then undiscounted.
for(i in 1:2){
  for(j in 1:NSAMPLES){
    dr = ifelse(i==1,pardf$discount[j],0)
    first0h = time_to_bpsv[j]+1
    bpsv_rd_py[[i]][j,first0h] = bpsv_upfront[j]/(1+dr)^(first0h-1) + inv_cost_per_year[j]/(1+dr)^(first0h-1)
    bpsv_rd_py[[i]][j,(first0h+1):15] = inv_cost_per_year[j]/(1+dr)^c(first0h:14)
  }
}

df = do.call(rbind,lapply(1:2, function(x) data.frame(years,d=c('Discounted','Undiscounted')[x],t(apply(t(apply(bpsv_rd_py[[x]]/1e3,1,cumsum)), 2, function(y)quantile(y,c(1,3)/4))))))

bpplot = ggplot(df)  +
  geom_ribbon(aes(x=years-1,ymin=(X25.),ymax=(X75.), fill=d),alpha=0.5) + 
  scale_x_continuous(n.breaks = 8)+# = seq(0,to=max(years),by=3)) +
  theme_bw(base_size = 16) + 
  labs(x='Year',y='Billion USD',colour='',fill='') +
  theme(legend.position = 'top')
ggsave(bpplot, filename='../../cepi_results/bpplot.png')



```

```{r}

## bringing together response costs

# R&D

# ssv_rd_plot 
bau_vals = ssv_rd_costsamples[[1]]
ssv_rd_diff = lapply( 2:length(ssv_rd_costsamples), function(x) ( ssv_rd_costsamples[[x]] - bau_vals)/1e3) 
ssv_rd_df = as.data.frame(t(sapply( ssv_rd_diff, function(x) quantile(x,iqrs)) ))
ssv_rd_df$timing = 'Per pandemic'
ssv_rd_df$category = 'R&D'
ssv_rd_df$type = missions

# bpsvresrd
bpsv_rd_res = as.data.frame(t(quantile(bpsvresrd/1e3,iqrs)))
bpsv_rd_res$timing = 'Per pandemic'
bpsv_rd_res$category = 'R&D'
bpsv_rd_res$type = 'BPSV'

# manufacturing

# ssv; take scenarios c(1,3,4) ; already in billions
bau_ref = ssv_proccost_undiscounted[[1]]
ssv_res_diff = lapply(c(3,4), function(x)ssv_proccost_undiscounted[[x]] - bau_ref)
ssv_res_df = as.data.frame(t(sapply(ssv_res_diff, function(x)quantile(x,iqrs))))
ssv_res_df$timing = 'Per pandemic'
ssv_res_df$category = 'Procurement'
ssv_res_df$type = capas

# bpsv; in billions

bpsv_proc_res = as.data.frame(t(quantile(bpsvproc/1e3,iqrs)))
bpsv_proc_res$timing = 'Per pandemic'
bpsv_proc_res$category = 'Procurement'
bpsv_proc_res$type = 'BPSV'



# delivery
# ssv; in billions
bau_del = ssv_delivery_costs[[1]]
ssv_del_diff = lapply(c(3,4,11),function(x)(ssv_delivery_costs[[x]]-bau_del)/1e3)
ssv_del_df = as.data.frame(t(sapply(ssv_del_diff,function(x)quantile(x,iqrs))))
ssv_del_df$timing = 'Per pandemic'
ssv_del_df$category = 'Delivery'
ssv_del_df$type = c(capas, 'Equality + Delivery')

# bpsv
# quantile(bpsv_del_cost/1e9,iqrs)

bpsv_del = as.data.frame(t(quantile(bpsv_del_cost/1e3,iqrs)))
bpsv_del$timing = 'Per pandemic'
bpsv_del$category = 'Delivery'
bpsv_del$type = 'BPSV'


datares = rbind(ssv_rd_df, bpsv_rd_res, ssv_res_df, bpsv_proc_res, bpsv_del, ssv_del_df)
datares$timing = factor(datares$timing,levels=timings,labels=timings)
datares$type = factor(datares$type,levels=types,labels=types)
# ggplot(datares) + theme_bw(base_size = 15) +
#   geom_hline(yintercept=0,colour='grey') +
#   geom_vline(xintercept=0,colour='grey') +
#   facet_wrap(~timing, scales='free') +
#   geom_bar(aes(x=type,y=`50%`,fill=category),stat = 'identity', position = 'dodge') +
#   geom_errorbar(position = position_dodge(width = 0.9), width=0.6,aes(x=type,group=category,ymin=`25%`,ymax=`75%`),show.legend = F) +
#       theme(axis.text.x = element_text(angle = 60,  hjust=1,vjust=1)) +
#   labs(x='',y='Billion USD',fill='',size='',title='Costs vs. BAU')

unitplot = rbind(prepdata,subset(datares,category!=''))
unitplot$category = factor(unitplot$category,levels=c('R&D','Procurement','Delivery'))

unitcosts =
  ggplot(unitplot) + theme_bw(base_size = 15) +
  geom_hline(yintercept=0,colour='grey') +
  geom_vline(xintercept=0,colour='grey') +
  facet_wrap(~timing, scales='free', space = "free_x") +
  geom_bar(aes(x=type,y=`50%`,fill=category),stat = 'identity', position = 'dodge') +
  geom_errorbar(position = position_dodge(width = 0.9), width=0.6,aes(x=type,group=category,ymin=`25%`,ymax=`75%`),show.legend = F) +
      theme(axis.text.x = element_text(angle = 60,  hjust=1,vjust=1)) +
  labs(x='',y='Billion USD',fill='',size='') +
  scale_y_continuous(n.breaks = 9) +
  theme(legend.position = 'right')


slot_map <- unitplot %>%
  distinct(timing, category) %>%
  group_by(timing) %>%
  arrange(category, .by_group = TRUE) %>%         # choose ordering; change if desired
  mutate(
    K = n(),
    slot_id = row_number(),                        # 1..K
    slot = slot_id - (K + 1) / 2                   # centered slots: K=3 -> -1,0,1 ; K=2 -> -0.5,0.5
  ) %>%
  ungroup()

df <- unitplot %>%
  distinct(timing, type, category, `25%`, `50%`, `75%`) %>%
  left_join(slot_map, by = c("timing", "category")) %>%
  group_by(timing) %>%
  mutate(type_id = as.integer(factor(type))) %>%
  group_by(timing, type) %>%
  mutate(
    k_present = n(),                               # how many categories for this type in this facet
    slot_use = if_else(k_present == 1, 0, slot)     # <-- center single-category groups
  ) %>%
  ungroup() %>%
  mutate(
    group_width = 0.9,
    x = type_id + (slot_use / K) * group_width,
    bar_w = group_width / K
  )

df$x = df$x  + 10*sapply(df$timing,function(x)which(c('Upfront','Per year','Per pandemic')==x))
df$type_id = df$type_id + 10*sapply(df$timing,function(x)which(c('Upfront','Per year','Per pandemic')==x))
# df$type_id[match(sort(unique(df$type_id)),df$type_id)]


unitcosts = ggplot(df) +
  theme_bw(base_size = 15) +
  geom_hline(yintercept = 0, colour = "grey") +
  facet_wrap(~timing, scales = "free", space = "free_x") +
  geom_col(aes(x = x, y = `50%`, fill = category),
           width = df$bar_w, show.legend = TRUE) +
  geom_errorbar(aes(x = x, ymin = `25%`, ymax = `75%`),
                width = .7*df$bar_w, show.legend = FALSE) +
  scale_x_continuous(
    breaks = sort((df$type_id)),
    labels = df$type[match(sort((df$type_id)),df$type_id)]
  ) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1),
        legend.position = "right") +
  labs(x = "", y = "Billion USD", fill = "") +
  scale_y_continuous(n.breaks = 9)

ggsave(unitcosts + labs(title=''), filename='../../cepi_results/unit.png',height=5,width=10)


```


```{r}

setDT(unitplot)

unitplot[,`Cost vs. BAU` := paste0('$',format_to_print2(`50%`) ,'$'),by=.(timing,category,type)]
unitplot[`25%`!=`75%`,`Cost vs. BAU` := 
           paste0('$',format_to_print2(`50%`) ,'$ ($',                              format_to_print2(`25%`),',',format_to_print2(`75%`),'$)'),by=.(timing,category,type)]

lg = 'Cost differences: investments vs. BAU, for different types of investment.'

setorder(unitplot,timing, category, type)

if (!knitr::is_html_output()) {
  cat(pander(unitplot[,-c(1:3)],caption=lg, missing = ""))
}

```




```{r}
## summing prep costs

# bpsv scenarios
# bpsv_rd_costsamples_dyd billions, discounted
# inv_cost in millions, discounted
# crs in billions
bpsv_ss = as.data.frame(t(quantile(bpsv_rd_costsamples_dyd + inv_cost/1e3 + crs[1] *rawpar$cost_capres*fifteen_yr_discount_weight,iqrs)))
bpsv_ss$Scenario = scennames[2]
bpsv_ss$Category = 'BPSV'
bpsv_ss$Capacity = crs[1]
bpsv_ss$DM = 365


capres_ss = as.data.frame(t(sapply(1:length(crs),function(x) quantile( crs[x] *rawpar$cost_capres*fifteen_yr_discount_weight,iqrs))))
capres_ss$Scenario = scennames[c(1,3:4)]
capres_ss$Category = '365DM'
capres_ss$Capacity = crs
capres_ss$DM = 365

# enabling_costs5 in millions, discounted
dm200_ss = as.data.frame(t(sapply(1:length(crs),function(x) quantile( crs[x] *rawpar$cost_capres*fifteen_yr_discount_weight + enabling_costs_list[[2]]/1e3,iqrs))))
dm200_ss$Scenario = scennames[5:7]
dm200_ss$Category = '200DM'
dm200_ss$Capacity = crs
dm200_ss$DM = 200

# enabling_costs15 in millions, discounted
dm100_ss = as.data.frame(t(sapply(1:length(crs),function(x) quantile( crs[x] *rawpar$cost_capres*fifteen_yr_discount_weight + enabling_costs_list[[3]]/1e3,iqrs))))
dm100_ss$Scenario = scennames[8:10]
dm100_ss$Category = '100DM'
dm100_ss$Capacity = crs
dm100_ss$DM = 100


preps = rbind(bpsv_ss, capres_ss, dm200_ss, dm100_ss)
preps = rbind(preps, capres_ss[1,])
preps$Scenario[11] = scennames[11]
preps$Category[11] = 'Equality + Delivery'
```


```{r}

cost_samples = readRDS('../results/cost_samples.Rds')
cost_diffs = list()
for (i in 1:3){
  cost_diffs[[i]] = cost_samples[[i]]
  for(j in 1:ncol(cost_diffs[[i]]))
    cost_diffs[[i]][,j] = cost_diffs[[i]][,j] - cost_samples[[i]][,1]
}
all_resp_costs <- all_prep_costs <- list()

for(s in 1:length(scennames)){
  all_prep_costs[[s]] = quantile(cost_diffs[[1]][,s]+cost_diffs[[2]][,s]*fifteen_yr_discount_weight,iqrs)/1e3
  all_resp_costs[[s]] = quantile(cost_diffs[[3]][,s]*at_yr_15,iqrs)/1e3
}

preps = as.data.frame(do.call(rbind,all_prep_costs))
preps$Scenario = scennames
preps$Category = c('365 days', 'BPSV', rep('365 days',2),rep(c('200 days','100 days'),each=3), 'Equality + Delivery')
preps$Capacity = crs[scen_capres]
preps$DM = dms[scen_dm]
preps$timing = 'Preparedness'

resps = as.data.frame(do.call(rbind,all_resp_costs))
resps$Scenario = scennames
resps$Category = preps$Category[match(resps$Scenario,preps$Scenario)]
resps$Capacity = crs[scen_capres]
resps$DM = dms[scen_dm]
resps$timing = 'Response'

melted = reshape2::melt(rbind(preps,resps),id.vars=c('Scenario','Category','Capacity','DM','timing'))
# melted$value = melted$value/113.8e3*100
toplot = reshape2::dcast(melted,formula=Scenario+Category+Capacity+DM~timing+variable)

toplot$Category = factor(toplot$Category,levels=c('BPSV', '365 days','200 days','100 days', 'Equality + Delivery'),labels=c('BPSV', '365 days','200 days','100 days', 'Equality + Delivery'))
toplot$Capacity = factor(toplot$Capacity,levels=c(0, 0.7, 2),labels=c('0.5 billion', '1.2 billion','2.5 billion'))

fifcosts = ggplot(toplot) +
  geom_hline(yintercept=0,colour='grey') +
  geom_vline(xintercept=0,colour='grey') +
  geom_abline(linetype=2,slope = -1,intercept = 0,
              colour='navyblue',show.legend = T) +
  geom_segment(
    aes(x = 0, xend = 0, y = 0, yend = 0, linetype = "Break even"),
    inherit.aes = FALSE, alpha = 0, show.legend = TRUE
  ) +
  geom_point(#show.legend = c('colour'=F),
               aes(x=`Preparedness_50%`,y=`Response_50%`,colour=Category,size=Capacity)) +
  geom_text(aes(x=`Preparedness_50%`+.8,y=`Response_50%`+2.5*(1-2*as.numeric(Scenario=='S10')),label=Scenario)) +
  geom_errorbar(show.legend=F,aes(x=`Preparedness_50%`,colour=Category,ymin=`Response_25%`,ymax=`Response_75%`), width = 0) +
  geom_errorbarh(show.legend=F,aes(y=`Response_50%`,colour=Category,xmin=`Preparedness_25%`,xmax=`Preparedness_75%`), width = 0) +
  labs(linetype='',x='Preparedness costs vs. BAU, billion USD',y='Response costs vs. BAU, billion USD',colour='',size='Capacity\nreservation',
       title='B) Preparedness vs. response (compared to BAU; billion USD)') +
  scale_linetype_manual(values = c("Break even" = "dashed")) +
  guides(
    linetype = guide_legend(
      override.aes = list(colour = "navyblue",alpha=1),
      title = NULL
    ),
    colour = guide_legend(title = NULL)#,
  ) +
  theme(legend.title = element_text(margin = margin(0,0))) +
  theme_bw(base_size = 15)

final = ggarrange(unitcosts, fifcosts, ncol=1)
ggsave(fifcosts + labs(title=''), filename='../../cepi_results/fifteen.png',height=6,width=10)

```


# SSV delivery


| Category | Reserved capacity | Private response (existing capacity)  | Private response (built capacity)  |
|---|---|---|---|
| Annual manufacturing volume | By scenario (0.5--2.5B)| 9B minus reserved volume | 6B |
| Facility transition start | 7 weeks before vaccine approval | 7 weeks before vaccine approval | 7 weeks before vaccine approval |
| Weeks to initial manufacturing | 12 | 12 (BPSV) or 30 (no BPSV) | 48 |
| Scale-up weeks to full capacity | 10 | 16 | 16 |

Table: Manufacturing response timeline assumptions 



 

<!-- | Weeks from transition start | 0-11 | 12-21 | 22-29 | 30-45  | 46-47 | 48-63 | 64+ | -->
<!-- |---|---|---|---|---|---|---|---| -->
<!-- | Reserved Capacity (%)  || Scaling from 0-100 | 100 | 100 | 100 | 100 | 100 |  -->
<!-- | Private Capacity (Existing; %)  || | | Scaling from 0-100 | 100 | 100 | 100 |  -->
<!-- | Private Capacity (Response; %)  | | | |  | | Scaling from 0-100 | 100 | -->



## Timing

We assume that manufacturers will begin producing at risk once they have confidence in vaccine approval. Thus, facility transition (the reorganising of manufacturing facilities in preparation for SSV dose production) begins $I_0=7$ weeks before vaccine approval, which in turn depends on R\&D investments. We have three levels in our scenarios, corresponding to SSVs available in 100 days, 200 days, and 365 days. The total weeks taken for vaccine approval can be written as follows:


```{r}
eqtext = 'W^{(\\zeta)} = \\sum_{i=0}^3 W_i^{(\\zeta)}'
printmath(eqtext)
```

for $\zeta\in\lbrace 365, 200, 100\rbrace$. These work out as `r time_to_approval[1]`, `r time_to_approval[2]`, and `r time_to_approval[3]` weeks, respectively. Thus "week 0" for manufacturing occurs `r time_to_approval[1]-7`, `r time_to_approval[2]-7`, and `r time_to_approval[3]-7` weeks, respectively, after the new pathogen has been sequenced. We denote this variable $w_s^{(0)}$:


```{r}
eqtext = 'w_s^{(0)} = W^{(\\zeta(s))} - I_0 = \\begin{cases}45 & s\\in\\lbrace 0, 1, 2, 3, 10\\rbrace \\\\ 
22 & s\\in\\lbrace 4, 5, 6\\rbrace \\\\ 
7 & s\\in\\lbrace 7, 8, 9\\rbrace \\end{cases}'
printmath(eqtext)
```

## Production



We estimate total global manufacturing potential to be $M_G=`r rawpar$man_glo`$ billion doses per year, and the amount currently in operation to be $M_C=`r rawpar$man_curr`$. The volume that is reserved, in billion doses, including the HIC-specific reservation of $A_3=`r rawpar$hic_cap_res`$ billion doses, depends on the scenarios as follows:

```{r}
eqtext = 'M_{R,s} = \\begin{cases}A_3 & s\\in\\lbrace 0, 1, 4, 7, 10\\rbrace \\\\ 
A_3 + 0.7 & s\\in\\lbrace 2, 5, 8\\rbrace \\\\ 
A_3 + 2 & s\\in\\lbrace 3, 6, 9\\rbrace \\end{cases}'
printmath(eqtext)
```

where $s=0$ denotes the BAU scenario. By definition, $M_{E,s} = M_C - M_{R,s}$ (existing unreserved capacity ("E") equals the total currently existing ("C") minus reserved capacity ("R")), and $M_B=M_G-M_C$ (newly built manufacturing ("B") equals the global total ("G") minus the existing capacity).

Then the number of doses, in billions, that are made from capacity $x\in \lbrace R, E, B\rbrace$ in week $w$ of scenario $s$ is:


```{r}
eqtext = 'Z_{x,s,w} = \\begin{cases}0 & w-w_s^{(0)} \\leq I_x \\\\ 
\\frac{1}{52}\\frac{w-w_s^{(0)}-I_x}{C_x}M_{x,s} & w-w_s^{(0)}\\in(I_x, I_x+C_x] \\\\ 
\\frac{1}{52}M_{x,s}  & w-w_s^{(0)}> I_x+C_x
\\end{cases}'
printmath(eqtext)
```

<!-- \frac{1}{52}M_{R,s}  & w\in[I_R+C_R, I_E)\\\\  -->
<!-- \frac{1}{52}\left(M_{R,s} + \frac{w-I_E+1}{C_E}M_{E,s}\right) & w\in[I_E, I_E+C_E)\\\\  -->
<!-- \frac{1}{52}\left(M_{R,s} + M_{E,s}\right)  & w\in[I_E+C_E, I_B)\\\\  -->
<!-- \frac{1}{52}\left(M_{R,s} + M_{E,s} + \frac{w-I_B+1}{C_B}M_{B}\right) & w\in[I_B, I_B+C_B)\\\\  -->

Here, $I_R = `r rawpar$weeks_init_res`$ is the number of weeks to initial manufacturing for reserved capacity, assuming streamlined technology transfer and process set up (around 2 months) and facility process validation (around 1 month) [@Kis2021a]. 
$C_R = `r rawpar$weeks_scale_res`$ is its number of weeks to scale up to full capacity, taking the lower end of the mRNA range assumed readiness from reserved capacity in @VaccinesEurope2024. 
$I_B = `r rawpar$weeks_init_bui`$ is the number of weeks to initial manufacturing for newly built capacity. This assumes early planning phase (around 2 months), longer facility improvement to build new infrastructure (around 5 months), tech transfer (around 2 months), and facility/process setup (around 2 months) [@Kis2021a]. This also aligns with realistic estimate from manufacturing stakeholder interviews [@Feddema2023].
$C_B = `r rawpar$weeks_scale_bui`$ is its number of weeks to scale up to full capacity, taking the middle of the range from @VaccinesEurope2024.

```{r}
eqtext = 'I_E = \\begin{cases}
 I_{E,1} \\; & \\; s=1 \\\\
I_{E,0}  \\; & \\; s\\neq 1
\\end{cases}'
printmath(eqtext)
```

where $I_{E,0} = `r rawpar$weeks_init_ex_nb`$ and $I_{E,1} = `r rawpar$weeks_init_ex_bp`$ are the number of weeks to initial manufacturing for existing and unreserved capacity (i.e. manufacturing conversion for SSVs is faster in the presence of BPSVs as we assume that they are manufactured using the same platform). In the absence of BPSV, we assume early planning phase (around 2 months), streamlined facility improvement because using existing infrastructure (around 1 month), tech transfer (around 2 months), and facility/process set up (around 2 months) [@Kis2021a].
$C_E = `r rawpar$weeks_scale_ex`$ is its number of weeks to scale up to full capacity, assumed to be the same as $C_B$. These timelines are shown in Table \@ref(tab:mantimeline).


| Weeks from transition start |	Reserved Capacity (%)	 | Existing Private Capacity (%)	| Response Private Capacity (%) |
|---|---|---|---|
| 0--11 |  |  |  |			
| 12--21 |	Scales from 0 to 100	 |  |  |				
| 22--29 |	100		|   |  |			
| 30--45 |	100 |	Scales from 0 to 100	  |  |
| 46--47 |	100 |	100	  |  |
| 48--63	 |100 |	100 |	Scales from 0 to 100 |
| 64+|	100 |	100 |	100 |

Table: (\#tab:mantimeline) Vaccine Production Timeline when there is no BPSV. When BPSV is also modelled, Existing Private Capacity scales from 0 to 100 in weeks 12--21.






```{r supply, fig.cap='Doses made available from manufacturing per scenario. Weeks are in reference to the sequencing of the pathogen.'}

t_supply <- as.data.frame(do.call(rbind,lapply(1:length(scennames),function(x)cbind(scen=scennames[x],doses=cumulative[[x]],week=1:NWEEKS))))

names(cumulative) = scennames

scens = setDT(data.frame(scen=scennames, capres = scen_capres, dm = scen_dm, bpsv = as.numeric(bpsv_scen)))
scens[,comb := paste0(capres,dm,bpsv)]
scens[,N:=.N,by=comb]
sames <- list()
for(un in unique(scens$comb)){
  subscen = subset(scens,comb==un)
  if(nrow(subscen)>1)
    sames[[length(sames)+1]] <- subscen$scen
}

ggplot(t_supply) + 
  geom_line(aes(x=as.numeric(week),y=as.numeric(doses),colour=factor(scen,levels=scennames))) +
  theme_bw(base_size = 15) +
  labs(x='Week',y='Cumulative doses manufactured, billions',colour='Scenario')
# ggplot(subset(t_supply,scen%in%scennames[2])) + 
#   geom_line(aes(x=as.numeric(week),y=as.numeric(doses),colour=factor(scen,levels=scennames))) +
#   theme_bw(base_size = 15) +
#   labs(x='Week',y='Cumulative doses manufactured, billions',colour='Scenario') +
#   scale_x_continuous(limits= c(55,65)) + 
#   scale_y_continuous(limits=c(0,1/2))
```

<!-- In Figure \@ref(fig:supply), the following scenarios have identical supply (because they have the same capacity reservations and R\&D investments): `r paste0(sames[[1]],collapse=' & ')`. -->

## Allocation


Denote the weekly allocated doses at week $w$ from capacity $x$ to income level $i$ $k_{s,x,i,w}$, and the cumulative number $K_{s,i,w}$, such that 


```{r}
eqtext = 'K_{s,i,w} = \\sum_{x\\in\\lbrace R,E,B\\rbrace}\\sum_{j=0}^w k_{s,x,i,j}.'
printmath(eqtext)
```




<!-- assuming vaccine wastage of $\delta = `r round(rawpar$vaccine_wastage,2)`$. -->

```{r}
eqtext = 'k_{s,R,i,w} = \\begin{cases}
\\left(\\frac{A_3}{M_{R,s}} + \\frac{M_{R,s}-A_3}{M_{R,s}}\\frac{N_{HIC}}{N_T}\\right)Z_{R,s,w}             & K_{s,\\text{HIC},w} <  L_{\\text{HIC}}\\;\\&\\; i=\\text{HIC} \\\\
\\frac{M_{R,s}-A_3}{M_{R,s}}\\frac{N_{i}}{N_T}Z_{R,s,w}                    & K_{s,\\text{HIC},w} < L_{\\text{HIC}}\\;\\&\\; i\\neq\\text{HIC}  \\\\
0 & K_{s,\\text{HIC},w} \\geq L_{\\text{HIC}} \\;\\&\\; i=\\text{HIC}\\\\
\\frac{N_{i}}{N_{UMIC}+N_{LMIC}+N_{LIC}}Z_{R,s,w} & K_{s,\\text{HIC},w} \\geq L_{\\text{HIC}} \\;\\&\\;  K_{s,\\text{UMIC},w} < L_{\\text{UMIC}} \\;\\&\\; i\\neq\\text{HIC}\\\\
0 & K_{s,\\text{UMIC},w} \\geq L_{\\text{UMIC}} \\;\\&\\; i=\\text{UMIC}\\\\
\\frac{N_{i}}{N_{LMIC}+N_{LIC}}Z_{R,s,w} & K_{s,\\text{UMIC},w} \\geq L_{\\text{UMIC}} \\;\\&\\;  K_{s,\\text{LMIC},w} < L_{\\text{LMIC}} \\;\\&\\; i\\notin  \\lbrace\\text{HIC},\\text{UMIC}\\rbrace\\\\
0 & K_{s,\\text{LMIC},w} \\geq L_{\\text{LMIC}} \\;\\&\\; i=\\text{LMIC}\\\\
Z_{R,s,w}             & K_{s,\\text{LMIC},w} \\geq L_{\\text{LMIC}} \\;\\&\\; K_{s,\\text{LIC},w} < L_{\\text{LIC}} \\;\\&\\; i=\\text{LIC}\\\\
0 & K_{s,\\text{LIC},w} \\geq L_{\\text{LIC}} 
\\end{cases}'
printmath(eqtext)
```

where $N_T = \sum_{i\in\lbrace\text{HIC,UMIC,LMIC,LIC}\rbrace}N_i$.

The logic of this reads as follows:

- $A_3=`r rawpar$hic_cap_res`$ billion doses per year from reserved capacity go exclusively to HIC, which is expressed as a fraction of the total reservation, $M_{R,s}$
- Any remaining reserved capacity doses are allocated according to population
- Once HIC reach their total demand, doses from reserved capacity are split proportional to population between UMIC, LMIC and LIC, and so on




For $x\in\lbrace E,B\rbrace$,


```{r}
eqtext = 'k_{s,x,i,w} = \\begin{cases}
Z_{x,s,w}            & K_{s,\\text{HIC},w} < L_{\\text{HIC}} \\;\\&\\; i=\\text{HIC} \\\\
0                     & K_{s,\\text{HIC},w} < L_{\\text{HIC}} \\;\\&\\; i\\neq\\text{HIC} \\\\
Z_{x,s,w}            & K_{s,\\text{HIC},w} \\geq L_{\\text{HIC}} \\;\\&\\; K_{s,\\text{UMIC},w} < L_{\\text{UMIC}} \\;\\&\\; i=\\text{UMIC} \\\\
0                     & K_{s,\\text{HIC},w} \\geq L_{\\text{HIC}} \\;\\&\\; K_{s,\\text{UMIC},w} < L_{\\text{UMIC}} \\;\\&\\; i\\neq\\text{UMIC} \\\\
Z_{x,s,w}            & K_{s,\\text{UMIC},w} \\geq L_{\\text{UMIC}} \\;\\&\\; K_{s,\\text{LMIC},w} < L_{\\text{LMIC}} \\;\\&\\; i=\\text{LMIC} \\\\
0                     & K_{s,\\text{UMIC},w} \\geq L_{\\text{UMIC}} \\;\\&\\; K_{s,\\text{LMIC},w} < L_{\\text{LMIC}} \\;\\&\\; i\\neq\\text{LMIC} \\\\
Z_{x,s,w}            & K_{s,\\text{LMIC},w} \\geq L_{\\text{LMIC}} \\;\\&\\; i=\\text{LIC} \\\\
0                     & K_{s,\\text{LMIC},w} \\geq L_{\\text{LMIC}} \\;\\&\\; i\\neq\\text{LIC} 
\\end{cases}'
printmath(eqtext)
```

The logic of this reads as follows:

- Until HIC demand is reached, all doses from unreserved capacity go to HIC. None go to UMIC, LMIC or LIC.
- Once HIC demand has been met and until UMIC demand is reached, all doses from unreserved capacity go to UMIC. None go to HIC, LMIC or LIC.
- Once HIC and UMIC demand have been met and until LMIC demand is reached, all doses from unreserved capacity go to LMIC. None go to HIC, UMIC or LIC.
- Once HIC, UMIC and LMIC demand have been met, all remaining doses from unreserved capacity go to LIC. None go to LMIC, UMIC or HIC.



Total supply of first-schedule doses in each year period is 


```{r}
eqtext = 'A_{x,s,y}^{(1)} = \\sum_i\\sum_{w\\in y}k_{s,x,i,w}.'
printmath(eqtext)
```

We assume, for every second dose of SSV, a booster will be given one year later for $N^{\text{(boost)}} = `r rawpar$n_boosters`$ years.

Thus 

```{r}
eqtext = 'A_{s,y}^{(2)} = \\begin{cases}
\\frac{1}{2}\\sum_{x}A_{x,s,y-1}^{(1)}            & y=2 \\\\
\\frac{1}{2}\\sum_{x}\\left(A_{x,s,y-1}^{(1)} + A_{x,s,y-2}^{(1)}\\right)            & y>2 
\\end{cases}'
printmath(eqtext)
```

and

```{r}
eqtext = 'A_{x,s,y}^{(2)} = \\min(A_{s,y}^{(2)}, M_R - A_{R,s,y}^{(1)})'
printmath(eqtext)
```

for $x=R$ and

```{r}
eqtext = 'A_{x,s,y}^{(2)} = \\max(A_{s,y}^{(2)} - A_{R,s,y}^{(2)}, 0)'
printmath(eqtext)
```

for $x\in\lbrace E, B \rbrace$.

Then

```{r}
eqtext = 'A_{x,s,y} = A_{x,s,y}^{(1)} + A_{x,s,y}^{(2)}.
(\\#eq:supply)'
if (knitr::is_html_output()) {
  cat(paste0("\\begin{equation}\n",gsub("\\\\(?!([a-z]|#))", "\\\\\\\\", eqtext, perl = TRUE),"\n\\end{equation}\n",collapse=''))
} else {
  cat(paste0("\\begin{equation}\n",eqtext,"\\end{equation}\n",collapse=''))
}
```




```{r procurement,fig.cap='Doses procured by country income level'}

ggplot(reshape2::melt(do.call(rbind,alloc),id.vars=c('Scenario','Week'))) +
  geom_line(aes(x=Week,colour=Scenario,y=value)) +
  facet_wrap(~variable) +
  theme_bw(base_size = 15) +
  labs(x='Week',y='Cumulative doses procured, billions',colour='Scenario')

```

## Delivery

<!-- **These values do not look correct** -->


Delivery is written as $h_{s,i,w}^{(j)}$ doses delivered in scenario $s$, income group $i$ and week $w$ for schedule $j$, which is $j=1$ for first-dose SSV, $j=2$ for second-dose SSV, and $j=2+k$ for $k=\lbrace 1,...,N^{(boost)}\rbrace$.

The second dose is prioritised over the first, and follows the first by four weeks, so 

```{r}
eqtext = 'h_{s,i,w}^{(2)} = h_{s,i,w-4}^{(1)}.'
printmath(eqtext)
```



Available doses are $K_{s,i,w}$, the cumulative doses allocated to income group $i$ by week $w$, minus doses given so far. First doses stop being given once $L_i/2$ is reached.

```{r,eval=T}
eqtext = 'h_{s,i,w}^{(1)} =
\\max\\left( 0,
\\min\\left(  K_{s,i,w} - h_{s,i,w}^{(2)} - \\sum_{j=1}^2 \\sum_{k=1}^{w-1} h_{s,i,k}^{(j)},
L_i/2 - \\sum_{k=1}^{w-1} h_{s,i,k}^{(1)}
\\right)
\\right)'
printmath(eqtext)
```


Booster doses are given for $j=2+k$ for $k=\lbrace 1,...,N^{(boost)}\rbrace$:

```{r}
eqtext = 'h_{s,i,w}^{(2+k)} = h_{s,i,w-52k}^{(2)}.'
printmath(eqtext)
```

Total doses are therefore

```{r}
eqtext = 'h_{s,i,w} = \\sum_{j=1}^{2+N^{(boost)}}h_{s,i,w}^{(j)}.'
printmath(eqtext)
```


```{r scendelivery,fig.cap='Cumulative vaccine coverage (second SSV dose) by country income level'}


second_dose_delivery = lapply(1:length(second_dose_delivery),function(x)cbind(second_dose_delivery[[x]],Scenario=scennames[x]))
ggplot(do.call(rbind,second_dose_delivery)) +
  geom_line(aes(x=Week*7,colour=Scenario,y=doses*100)) +
  facet_wrap(~factor(il,levels=INCOMELEVELS)) +
  theme_bw(base_size = 15) +
  scale_x_continuous(limits=c(0,1300)) +
  labs(x='Day',y='Second-dose coverage, percent population',colour='Scenario')


```


The input to the impact model is $h_{s,i,w}^{(2)}$, i.e. doses delivered per week, per scenario, per income level. The weekly value is transformed to a daily value by dividing by 7, i.e. every day in a given week sees the same number of doses distributed. The impact model combines LIC and LMIC into a single category, LLMIC. For this, a population-weighted average is taken of $h_{s,\text{LIC},w}^{(2)}$ and $h_{s,\text{LMIC},w}^{(2)}$. The same applies to BPSV delivery.


# BPSV delivery

## Timing

The duration of the Phase 3 trial is $W_3^{(365)} = `r rawpar$weeks_P3_365`$ weeks. The time to manufacturing transition is $I_R = `r rawpar$weeks_init_res`$ weeks, and the time to manufacturing scale-up $C_R = `r rawpar$weeks_scale_res`$ weeks; these are the same as the reserved-capacity times for SSV.


Facility transition occurs in week 1. Thus manufacturing begins in week $1+I_R = `r 1+rawpar$weeks_init_res`$ and dose distribution begins in week $1+W_3^{(365)} = `r 1+rawpar$weeks_P3_365`$.

## Production

The number of doses, in billions, that are made in week $w$ is:


```{r}
eqtext = 'Z_w = \\begin{cases}0 & w < I_R \\\\ 
\\frac{1}{52}\\frac{w-I_x+1}{C_x}M_{x,s} & w\\in[I_R, I_R+C_R) \\\\ 
\\frac{1}{52}M_{x,s}  & w-1\\geq I_R+C_xR
\\end{cases}'
printmath(eqtext)
```




```{r bpsvsupply, fig.cap='BPSV doses made available from manufacturing. Weeks are in reference to the sequencing of the pathogen.'}

ggplot(b_supply) + 
  geom_line(aes(x=as.numeric(week),y=as.numeric(doses))) +
  theme_bw(base_size = 15) +
  labs(x='Week',y='Cumulative doses manufactured, billions',colour='Scenario')

```


## Allocation

Doses are all allocated in proportion to eligible populations.



```{r bpsvprocurement,fig.cap='BPSV doses procured by country income level'}

ggplot(reshape2::melt(b_alloc,id.vars=c('Week'))) +
  geom_line(aes(x=Week,colour=variable,y=value)) +
  theme_bw(base_size = 15) +
  labs(x='Week',y='Cumulative doses procured, billions',colour='Income level')

```

## Delivery

<!-- **These values do not look correct** -->


```{r bpsvdeliveryplot,fig.cap='BPSV vaccine coverage by country income level'}

ggplot(reshape2::melt(bpsv_dose_delivery,id.vars=c('Week'))) +
  geom_line(aes(x=Week*7,colour=factor(variable,levels=INCOMELEVELS,labels=INCOMELEVELS),y=value*100)) +
  theme_bw(base_size = 15) +
  labs(x='Day',y='BPSV coverage, percent population',colour='Income level')


```



# Parameter samples

```{r parsamples}

for(i in 1:length(params))
  base::print(
    ggplot() + 
      geom_histogram(aes(x=params[[i]])) + 
      theme_bw(base_size = 13) + 
      labs(x='',y='',title=names(params)[i])
    )

```


# Contributors

Model: Peter Windus, Andy Torkelson

Data: Peter Windus, Andy Torkelson, Damian Walker

Documentation: Peter Windus, Andy Torkelson, Rob Johnson

R code: Rob Johnson


# References
